---
title: "TiNTS epidemiological analysis: Cox models"
format: html
execute:
  echo: true
  warning: true
  message: true
params:
  project_root: "."
  strict_reference_check: true
editor: source
---

This document reconstructs the final covariates and reproduces the first-event Cox proportional-hazards model and Andersen–Gill recurrent-event model reported in the manuscript. It includes fail-fast checks against the archived manuscript estimates.

## Setup

```{r}
options(stringsAsFactors = FALSE, scipen = 999)

required_packages <- c("dplyr", "readr", "stringr", "tibble", "survival")

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0L) {
  stop(
    "Install the following packages before rendering: ",
    paste(missing_packages, collapse = ", "),
    call. = FALSE
  )
}

suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
  library(stringr)
  library(tibble)
  library(survival)
})

# Locate the TiNTS project root automatically.  The analysis script lives in
# <project>/scripts, whereas data_raw/, data_clean/, and rds/ live one level
# above it.  Quarto/RStudio may render with the working directory set either to
# the project root or to scripts/, so do not assume that getwd() is the root.
required_relative_files <- c(
  file.path("data_raw", "CRF_A.csv"),
  file.path("data_raw", "CRF_B_MCA.csv"),
  file.path("data_clean", "wealth_index_by_household.csv"),
  file.path("rds", "episodes_df.rds"),
  file.path("rds", "followup_df.rds")
)

normalise_existing_dir <- function(x) {
  if (is.null(x) || length(x) == 0L || is.na(x) || !nzchar(x)) return(NULL)
  normalizePath(x, winslash = "/", mustWork = FALSE)
}

ancestor_dirs <- function(path, n = 8L) {
  path <- normalise_existing_dir(path)
  if (is.null(path)) return(character())
  out <- path
  for (i in seq_len(n)) {
    parent <- dirname(out[length(out)])
    if (identical(parent, out[length(out)])) break
    out <- c(out, parent)
  }
  unique(out)
}

qmd_dir <- tryCatch({
  input <- knitr::current_input(dir = TRUE)
  if (is.null(input) || !nzchar(input)) NULL else dirname(normalise_existing_dir(input))
}, error = function(e) NULL)

quarto_project_dir <- normalise_existing_dir(Sys.getenv("QUARTO_PROJECT_DIR", unset = ""))
explicit_root <- normalise_existing_dir(params$project_root)

root_candidates <- unique(c(
  ancestor_dirs(explicit_root),
  ancestor_dirs(getwd()),
  ancestor_dirs(qmd_dir),
  ancestor_dirs(quarto_project_dir)
))
root_candidates <- root_candidates[nzchar(root_candidates)]

is_tints_root <- function(path) {
  all(file.exists(file.path(path, required_relative_files)))
}

root_ok <- vapply(root_candidates, is_tints_root, logical(1))

if (!any(root_ok)) {
  searched <- if (length(root_candidates)) {
    paste0("- ", root_candidates, collapse = "\n")
  } else {
    "- <no candidate directories resolved>"
  }
  stop(
    "Could not locate the TiNTS project root automatically.\n",
    "I looked for these five required files beneath each candidate root:\n- ",
    paste(required_relative_files, collapse = "\n- "),
    "\n\nDirectories searched:\n", searched,
    "\n\nIf the project has moved, set params$project_root once to the folder that contains data_raw/, data_clean/, and rds/.",
    call. = FALSE
  )
}

root <- normalizePath(
  root_candidates[which(root_ok)[1]],
  winslash = "/",
  mustWork = TRUE
)

message("TiNTS project root: ", root)

project_file <- function(...) file.path(root, ...)

input_files <- c(
  crf_a    = project_file("data_raw", "CRF_A.csv"),
  crf_b    = project_file("data_raw", "CRF_B_MCA.csv"),
  wealth   = project_file("data_clean", "wealth_index_by_household.csv"),
  episodes = project_file("rds", "episodes_df.rds"),
  followup = project_file("rds", "followup_df.rds")
)

# This is deliberately retained as a second, explicit guard: if the root was
# found but one of the required files subsequently disappears, fail clearly.
missing_files <- input_files[!file.exists(input_files)]
if (length(missing_files) > 0L) {
  stop(
    "Required input file(s) not found beneath detected project root ", root, ":\n",
    paste0("- ", missing_files, collapse = "\n"),
    call. = FALSE
  )
}



assert_columns <- function(data, columns, object_name) {
  absent <- setdiff(columns, names(data))
  if (length(absent) > 0L) {
    stop(
      object_name, " is missing required column(s): ",
      paste(absent, collapse = ", "),
      call. = FALSE
    )
  }
  invisible(data)
}

clean_text <- function(x) stringr::str_squish(trimws(as.character(x)))

yes_no <- function(x) {
  x <- tolower(clean_text(x))
  dplyr::case_when(
    x == "yes" ~ 1L,
    x == "no"  ~ 0L,
    TRUE        ~ NA_integer_
  )
}

safe_min_date <- function(x) {
  if (all(is.na(x))) as.Date(NA) else min(x, na.rm = TRUE)
}

safe_max_date <- function(x) {
  if (all(is.na(x))) as.Date(NA) else max(x, na.rm = TRUE)
}
```

## Final-model covariates

```{r}
A_raw       <- readr::read_csv(input_files[["crf_a"]], show_col_types = FALSE)
B_raw       <- readr::read_csv(input_files[["crf_b"]], show_col_types = FALSE)
hh_wealth   <- readr::read_csv(input_files[["wealth"]], show_col_types = FALSE)
episodes_df <- readr::read_rds(input_files[["episodes"]])
followup_df <- readr::read_rds(input_files[["followup"]])

assert_columns(
  A_raw,
  c(
    "hhid", "hh_sourceofwater", "hh_watertreated",
    "before_eatn", "before_prepfood", "after_toilet", "lookn_dirty",
    "perishable_food", "hh_washing"
  ),
  "CRF_A.csv"
)

assert_columns(
  B_raw,
  c(
    "pid", "hhid", "age_yr", "sex", "dys_leav_day", "cls_contactnum",
    "eat_atcafe", "eat_meat", "eat_meatfreq", "eat_fish", "eat_fishfreq",
    "hrs_spentnhouse", "leav_ndirnd"
  ),
  "CRF_B_MCA.csv"
)

assert_columns(hh_wealth, c("hhid", "wealth_index"), "wealth_index_by_household.csv")
assert_columns(
  episodes_df,
  c("pid", "hhid", "episode_order", "onset_lower", "onset_upper", "end_upper"),
  "episodes_df.rds"
)
assert_columns(followup_df, c("pid", "hhid", "first_obs", "last_obs"), "followup_df.rds")

# Only the 12 elastic-net-selected predictor groups and the three
# prespecified confounders are reconstructed here. These transformations
# reproduce the final-model coding used for the manuscript analysis.

drinking_sources_known <- c(
  "Public tap/Standpipe",
  "Piped into dwelling",
  "Piped outside dwelling",
  "Tube well / borehole",
  "Tube well/borehole",
  "Borehole",
  "Unprotected well / spring",
  "Unprotected well/spring"
)

food_storage_known <- c(
  "Container",
  "Dark cupboard",
  "Eaten same day",
  "Food is left on surfaces",
  "Refrigerator"
)

A_covariates <- A_raw %>%
  transmute(
    hhid = as.character(hhid),
    drinking_source = clean_text(hh_sourceofwater),
    utensil_source = clean_text(hh_washing),
    food_storage = clean_text(perishable_food),
    water_treated = yes_no(hh_watertreated),
    hand_before_eating = yes_no(before_eatn),
    hand_before_food = yes_no(before_prepfood),
    hand_after_toilet = yes_no(after_toilet),
    hand_when_dirty = yes_no(lookn_dirty)
  ) %>%
  mutate(
    drinking_water_binary = factor(
      case_when(
        is.na(drinking_source) ~ NA_character_,
        !drinking_source %in% drinking_sources_known ~ NA_character_,
        drinking_source %in% c("Piped into dwelling", "Piped outside dwelling") ~ "piped",
        TRUE ~ "non_piped"
      ),
      levels = c("non_piped", "piped")
    ),
    utensils_washed_binary = factor(
      case_when(
        is.na(utensil_source) ~ NA_character_,
        utensil_source %in% c("Piped into dwelling", "Piped outside dwelling") ~ "piped",
        TRUE ~ "non_piped"
      ),
      levels = c("non_piped", "piped")
    ),
    food_storage_binary = factor(
      case_when(
        is.na(food_storage) ~ NA_character_,
        !food_storage %in% food_storage_known ~ NA_character_,
        food_storage == "Refrigerator" ~ "refrigerated",
        TRUE ~ "unrefrigerated"
      ),
      levels = c("unrefrigerated", "refrigerated")
    ),
    hand_index = (
      hand_before_eating + hand_before_food +
        hand_after_toilet + hand_when_dirty
    ) / 4
  ) %>%
  select(
    hhid,
    drinking_water_binary,
    utensils_washed_binary,
    food_storage_binary,
    water_treated,
    hand_index
  )

hours_midpoint <- c(
  "Less than 5 hours"  = 2.5,
  "5-8 hours"          = 6.5,
  "9-13 hours"         = 11,
  "14-19 hours"        = 16.5,
  "More than 20 hours" = 22
)

B_covariates <- B_raw %>%
  transmute(
    pid = as.character(pid),
    hhid = as.character(hhid),
    age_yr = as.numeric(age_yr),
    sex = as.character(sex),
    days_out_day = as.integer(dys_leav_day),
    cls_contactnum = as.integer(cls_contactnum),
    eat_atcafe = yes_no(eat_atcafe),
    leave_ndirande = clean_text(leav_ndirnd),
    meat_eaten = clean_text(eat_meat),
    meat_frequency_reported = clean_text(eat_meatfreq),
    fish_eaten = clean_text(eat_fish),
    fish_frequency_reported = clean_text(eat_fishfreq),
    hours_category = clean_text(hrs_spentnhouse)
  ) %>%
  mutate(
    ndir_leave_binary = factor(
      case_when(
        is.na(leave_ndirande) ~ NA_character_,
        leave_ndirande %in% c("At least once a week", "Every day") ~ "regularly",
        TRUE ~ "rarely"
      ),
      levels = c("rarely", "regularly")
    ),
    meat_frequency = case_when(
      meat_eaten == "No" ~ "Never",
      TRUE ~ meat_frequency_reported
    ),
    meat_freq_binary = factor(
      case_when(
        is.na(meat_frequency) ~ NA_character_,
        meat_frequency %in% c("At least once a month", "Every day") ~ "regularly",
        TRUE ~ "rarely"
      ),
      levels = c("rarely", "regularly")
    ),
    fish_frequency = case_when(
      fish_eaten == "No" ~ "Never",
      TRUE ~ fish_frequency_reported
    ),
    fish_freq_binary = factor(
      case_when(
        is.na(fish_frequency) ~ NA_character_,
        fish_frequency %in% c("At least once a month", "Every day") ~ "regularly",
        TRUE ~ "rarely"
      ),
      levels = c("rarely", "regularly")
    ),
    hours_in_house_num = unname(hours_midpoint[hours_category])
  ) %>%
  select(
    pid, hhid, age_yr, sex,
    days_out_day, cls_contactnum, eat_atcafe,
    ndir_leave_binary, meat_freq_binary, fish_freq_binary,
    hours_in_house_num
  )

wealth_covariate <- hh_wealth %>%
  transmute(
    hhid = as.character(hhid),
    wealth_index_z = as.numeric(wealth_index)
  )

if (anyDuplicated(A_covariates$hhid)) {
  stop("CRF A contains duplicate household IDs.", call. = FALSE)
}
if (anyDuplicated(B_covariates$pid)) {
  stop("CRF B contains duplicate participant IDs.", call. = FALSE)
}
if (anyDuplicated(wealth_covariate$hhid)) {
  stop("The wealth-index file contains duplicate household IDs.", call. = FALSE)
}

covariates <- B_covariates %>%
  left_join(A_covariates, by = "hhid") %>%
  left_join(wealth_covariate, by = "hhid")
```

## First-event Cox model

```{r}
episodes_df <- episodes_df %>%
  mutate(
    pid = as.character(pid),
    hhid = as.character(hhid),
    episode_order = as.integer(episode_order),
    onset_lower = as.Date(onset_lower),
    onset_upper = as.Date(onset_upper),
    end_upper = as.Date(end_upper)
  )

followup_df <- followup_df %>%
  mutate(
    pid = as.character(pid),
    hhid = as.character(hhid),
    first_obs = as.Date(first_obs),
    last_obs = as.Date(last_obs)
  )

first_events <- episodes_df %>%
  filter(
    episode_order == 1L,
    !is.na(onset_lower),
    !is.na(onset_upper)
  ) %>%
  transmute(
    pid, hhid,
    tstart = onset_lower,
    tstop = onset_upper,
    event = 1L
  )

first_censors <- followup_df %>%
  anti_join(first_events %>% distinct(pid), by = "pid") %>%
  transmute(
    pid, hhid,
    tstart = first_obs,
    tstop = last_obs,
    event = 0L
  )

selection_df <- bind_rows(first_events, first_censors) %>%
  left_join(covariates, by = c("pid", "hhid")) %>%
  mutate(
    tstart = as.Date(tstart),
    tstop = as.Date(tstop)
  )

study_start <- min(selection_df$tstart, na.rm = TRUE)

selection_df <- selection_df %>%
  mutate(
    tstart_num = as.numeric(difftime(tstart, study_start, units = "days")),
    tstop_num = as.numeric(difftime(tstop, study_start, units = "days"))
  )

final_model_terms <- c(
  "drinking_water_binary",
  "utensils_washed_binary",
  "food_storage_binary",
  "ndir_leave_binary",
  "meat_freq_binary",
  "fish_freq_binary",
  "days_out_day",
  "cls_contactnum",
  "eat_atcafe",
  "water_treated",
  "hand_index",
  "hours_in_house_num",
  "age_yr",
  "sex",
  "wealth_index_z"
)

first_formula <- as.formula(
  paste(
    "Surv(tstart_num, tstop_num, event) ~",
    paste(final_model_terms, collapse = " + ")
  )
)

fit_first <- survival::coxph(
  first_formula,
  data = selection_df,
  ties = "efron",
  cluster = hhid,
  na.action = na.omit,
  model = TRUE,
  x = TRUE,
  y = TRUE
)
```

## Andersen–Gill recurrent-event model

```{r}
# This reproduces the recurrent-event interval construction in the source
# workflow: total time from each participant's episode-derived baseline, with
# robust standard errors clustered by participant.

person_dates <- episodes_df %>%
  mutate(
    start_proxy = coalesce(onset_lower, onset_upper),
    end_proxy = coalesce(end_upper, onset_upper)
  ) %>%
  group_by(pid, hhid) %>%
  summarise(
    baseline_date = safe_min_date(start_proxy),
    last_followup_date = safe_max_date(end_proxy),
    .groups = "drop"
  )

ag_events <- episodes_df %>%
  left_join(person_dates, by = c("pid", "hhid")) %>%
  arrange(pid, episode_order) %>%
  group_by(pid) %>%
  mutate(
    event_date = onset_upper,
    episode_end = coalesce(end_upper, onset_upper),
    start_date = if_else(
      episode_order == 1L,
      baseline_date,
      lag(episode_end)
    ),
    stop_date = event_date,
    event = 1L
  ) %>%
  ungroup() %>%
  mutate(
    tstart_num = as.numeric(difftime(start_date, baseline_date, units = "days")),
    tstop_num = as.numeric(difftime(stop_date, baseline_date, units = "days"))
  ) %>%
  select(pid, hhid, tstart_num, tstop_num, event)

censor_after_last <- episodes_df %>%
  left_join(person_dates, by = c("pid", "hhid")) %>%
  mutate(episode_end = coalesce(end_upper, onset_upper)) %>%
  group_by(pid, hhid, baseline_date, last_followup_date) %>%
  summarise(
    last_episode_end = safe_max_date(episode_end),
    .groups = "drop"
  ) %>%
  filter(
    !is.na(last_followup_date),
    !is.na(last_episode_end),
    last_followup_date > last_episode_end
  ) %>%
  transmute(
    pid, hhid,
    tstart_num = as.numeric(difftime(last_episode_end, baseline_date, units = "days")),
    tstop_num = as.numeric(difftime(last_followup_date, baseline_date, units = "days")),
    event = 0L
  )

never_event_pids <- setdiff(
  unique(selection_df$pid),
  unique(episodes_df$pid)
)

never_event_censors <- selection_df %>%
  filter(pid %in% never_event_pids) %>%
  group_by(pid, hhid) %>%
  summarise(
    tstart_num = min(tstart_num, na.rm = TRUE),
    tstop_num = max(tstop_num, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(
    is.finite(tstart_num),
    is.finite(tstop_num),
    tstop_num > tstart_num
  ) %>%
  mutate(event = 0L)

ag_intervals <- bind_rows(
  ag_events,
  censor_after_last,
  never_event_censors
) %>%
  filter(
    is.finite(tstart_num),
    is.finite(tstop_num),
    tstop_num > tstart_num
  )

analysis_ag <- ag_intervals %>%
  left_join(
    selection_df %>%
      distinct(pid, .keep_all = TRUE) %>%
      select(pid, hhid, all_of(final_model_terms)),
    by = c("pid", "hhid")
  )

ag_formula <- as.formula(
  paste(
    "Surv(tstart_num, tstop_num, event) ~",
    paste(final_model_terms, collapse = " + ")
  )
)

fit_ag <- survival::coxph(
  ag_formula,
  data = analysis_ag,
  ties = "efron",
  cluster = pid,
  na.action = na.omit,
  model = TRUE,
  x = TRUE,
  y = TRUE
)
```

## Extract and audit model estimates

```{r}
term_dictionary <- tribble(
  ~term_id, ~prefix,
  "drinking_water_binary",  "drinking_water_binary",
  "utensils_washed_binary", "utensils_washed_binary",
  "food_storage_binary",    "food_storage_binary",
  "ndir_leave_binary",      "ndir_leave_binary",
  "meat_freq_binary",       "meat_freq_binary",
  "fish_freq_binary",       "fish_freq_binary",
  "days_out_day",           "days_out_day",
  "cls_contactnum",         "cls_contactnum",
  "eat_atcafe",             "eat_atcafe",
  "water_treated",          "water_treated",
  "hand_index",             "hand_index",
  "hours_in_house_num",     "hours_in_house_num",
  "age_yr",                 "age_yr",
  "sex",                    "sex",
  "wealth_index_z",         "wealth_index_z"
)

map_model_term <- function(raw_term) {
  hits <- term_dictionary$term_id[
    startsWith(raw_term, term_dictionary$prefix)
  ]
  if (length(hits) != 1L) {
    stop("Could not uniquely map model term: ", raw_term, call. = FALSE)
  }
  hits[[1]]
}

extract_clustered_cox <- function(fit, model_name) {
  coefficient_table <- as.data.frame(summary(fit)$coefficients)

  if (!"robust se" %in% names(coefficient_table)) {
    stop(model_name, " did not return robust standard errors.", call. = FALSE)
  }

  p_column <- grep("^Pr\\(", names(coefficient_table), value = TRUE)[1]

  tibble(
    model = model_name,
    raw_term = rownames(coefficient_table),
    beta = coefficient_table[["coef"]],
    robust_se = coefficient_table[["robust se"]],
    p_value = coefficient_table[[p_column]]
  ) %>%
    mutate(
      term_id = vapply(raw_term, map_model_term, character(1)),
      hr = exp(beta),
      lower = exp(beta - 1.96 * robust_se),
      upper = exp(beta + 1.96 * robust_se)
    ) %>%
    select(model, term_id, hr, lower, upper, p_value)
}

model_results <- bind_rows(
  extract_clustered_cox(fit_first, "First event"),
  extract_clustered_cox(fit_ag, "First and recurrent events")
)

# Reference values reported in the manuscript and supplementary tables.
# The refitted results must reproduce these values to two decimal places.
reference_values <- tribble(
  ~model, ~term_id, ~ref_hr, ~ref_lower, ~ref_upper,
  "First event", "drinking_water_binary",  1.52, 0.81, 2.83,
  "First event", "utensils_washed_binary", 0.32, 0.13, 0.76,
  "First event", "water_treated",           1.56, 0.82, 3.00,
  "First event", "hand_index",              1.11, 0.36, 3.37,
  "First event", "food_storage_binary",     1.18, 0.48, 2.92,
  "First event", "meat_freq_binary",        1.54, 0.99, 2.41,
  "First event", "fish_freq_binary",        0.68, 0.39, 1.18,
  "First event", "eat_atcafe",              0.25, 0.07, 0.84,
  "First event", "ndir_leave_binary",       0.61, 0.30, 1.23,
  "First event", "days_out_day",            1.14, 1.03, 1.26,
  "First event", "cls_contactnum",          1.02, 1.00, 1.03,
  "First event", "hours_in_house_num",      0.95, 0.91, 1.01,
  "First event", "age_yr",                  1.00, 0.99, 1.02,
  "First event", "sex",                     1.26, 0.78, 2.04,
  "First event", "wealth_index_z",          1.23, 0.93, 1.64,
  "First and recurrent events", "drinking_water_binary",  0.78, 0.56, 1.10,
  "First and recurrent events", "utensils_washed_binary", 0.78, 0.43, 1.43,
  "First and recurrent events", "water_treated",           1.21, 0.84, 1.74,
  "First and recurrent events", "hand_index",              0.29, 0.14, 0.60,
  "First and recurrent events", "food_storage_binary",     1.47, 0.89, 2.42,
  "First and recurrent events", "meat_freq_binary",        0.81, 0.61, 1.08,
  "First and recurrent events", "fish_freq_binary",        0.59, 0.41, 0.84,
  "First and recurrent events", "eat_atcafe",              0.42, 0.21, 0.85,
  "First and recurrent events", "ndir_leave_binary",       1.33, 0.88, 2.01,
  "First and recurrent events", "days_out_day",            1.10, 1.04, 1.16,
  "First and recurrent events", "cls_contactnum",          1.00, 0.99, 1.01,
  "First and recurrent events", "hours_in_house_num",      1.00, 0.96, 1.03,
  "First and recurrent events", "age_yr",                  1.00, 0.99, 1.01,
  "First and recurrent events", "sex",                     1.24, 0.91, 1.71,
  "First and recurrent events", "wealth_index_z",          1.08, 0.89, 1.30
)

validation <- full_join(
  model_results,
  reference_values,
  by = c("model", "term_id")
) %>%
  mutate(
    model_hr = sprintf("%.2f", hr),
    model_lower = sprintf("%.2f", lower),
    model_upper = sprintf("%.2f", upper),
    reference_hr = sprintf("%.2f", ref_hr),
    reference_lower = sprintf("%.2f", ref_lower),
    reference_upper = sprintf("%.2f", ref_upper),
    matches = !is.na(hr) & !is.na(ref_hr) &
      model_hr == reference_hr &
      model_lower == reference_lower &
      model_upper == reference_upper
  )

first_event_count <- sum(fit_first$y[, 3])
first_person_days <- sum(fit_first$y[, 2] - fit_first$y[, 1])
ag_event_count <- sum(fit_ag$y[, 3])

count_mismatch <-
  !isTRUE(all.equal(as.numeric(first_event_count), 125)) ||
  !isTRUE(all.equal(as.numeric(first_person_days), 6957)) ||
  !isTRUE(all.equal(as.numeric(ag_event_count), 166))

estimate_mismatch <- validation %>% filter(!matches | is.na(matches))

if (nrow(model_results) != 30L) {
  count_mismatch <- TRUE
}

if (nrow(estimate_mismatch) > 0L || count_mismatch) {
  mismatch_text <- capture.output(
    print(
      estimate_mismatch %>%
        select(
          model, term_id,
          model_hr, model_lower, model_upper,
          reference_hr, reference_lower, reference_upper
        ),
      n = Inf
    )
  )

  audit_message <- paste0(
    "The refitted models do not reproduce the archived manuscript reference values.\n",
    "First-event count/person-days: ", first_event_count, "/", first_person_days,
    " (expected 125/6957). Recurrent-model events: ", ag_event_count,
    " (expected 166). Model rows: ", nrow(model_results), " (expected 30).\n\n",
    paste(mismatch_text, collapse = "\n"),
    "\n\nResolve the data/model discrepancy before using these results."
  )

  if (isTRUE(params$strict_reference_check)) {
    stop(audit_message, call. = FALSE)
  } else {
    warning(audit_message, call. = FALSE)
  }
}

dir.create(project_file("outputs", "reproducibility"), recursive = TRUE, showWarnings = FALSE)
readr::write_csv(
  model_results %>% filter(model == "First event"),
  project_file("outputs", "reproducibility", "cox_first_event_results.csv")
)
readr::write_csv(
  model_results %>% filter(model == "First and recurrent events"),
  project_file("outputs", "reproducibility", "cox_andersen_gill_results.csv")
)
readr::write_csv(
  validation,
  project_file("outputs", "reproducibility", "cox_reference_validation.csv")
)
```
