---
title: "Bayesian Model A: individual hidden Markov model"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

# Overview

Model A is a two-state susceptible–infected hidden Markov model for daily PCR observations. A sampled day is represented either by no amplification or by an observed Ct value. State-specific amplification probabilities and Gaussian Ct distributions distinguish the latent susceptible and infected states, while the latent process follows a daily SIS transition model.

The workflow reads the daily-grid modelling input, compiles the accompanying Stan program, fits the model, and generates posterior summaries and derived epidemiological quantities. Priors and sampler settings are specified explicitly below.

## i. The model's premise:

On each given day, the participant is in one of two latent biological states

-   S = not infected / not shedding

-   I - infected / shedding

The QUESTION the model is asking is: *"if I knew which biological state the participant was in on that day, how likely would each type of PCR outcome be?"*

## ii. The model's OBSERVATION:

The OBSERVATION is:

Either:

1.  "no Ct" (no amplification / not detected)

Or:

2.  a numeric Ct value.

This is not the same as what we had previously, where the observation was a numeric Ct value that could be present or absent... this is a categorical-numeric hybrid observation: either *nothing* was detected or a *number* was detected.

We use the "empty set" (in latex \varnothing) notation for this: $\varnothing$

So when we see $\varnothing$ it means "no Ct value was observed". For example:

$$y_t = \varnothing$$

means "On day $t$, the PCR result was no Ct"

The mathematical notation for encoding the observation is as follows:

$$
y_t \in \{\varnothing\} \cup \mathbb{R}
$$

This means:

The observation at time $t$

$$
y_t
$$

Is an element of

$$
\in
$$

The empty set

$$
\{\varnothing\}
$$

or

$$
\cup
$$

a real number

$$
\mathbb{R}
$$

Three examples to illustrate this:

| Day $t$ | PCR machine output | $y_t$                       |
|---------|--------------------|-----------------------------|
| 2       | Ct = 22.6          | $y_2 = 22.6 \in \mathbb{R}$ |
| 4       | PCR negative       | $y_4 = \varnothing$         |
| 6       | Ct = 37.9          | $y_2 = 37.9 \in \mathbb{R}$ |

## iii. The latent state

Let:

$$
Z_{i,t} \in \{S, I\}
$$

-   $S$ : not infected

-   $I$: infected

## iv. biological statement

For each state $\{S, I\}$ :

1.  PCR may or may not detect a TTR target DNA amplification
2.  Only if amplification occurs do we see a Ct value
3.  The Ct value has a distribution (depending on the state).

This means emissions are a two-stage process (detection / Ct value)

## v. Detection probabilities

-   $\phi_I$: sensitivity

-   $\phi_S$: false positive probability

$$
\phi_I = P(D_t = 1 \mid Z_t = I)
$$

$$
\phi_S = P(D_t = 1 \mid Z_t = S)
$$

Where $D_t$ is "\[PCR\] detection at time t"

and $Z_t$ is "state at time t"

## vi. Probability of no Ct

If no DNA amplification is detected, the observed outcome is $\varnothing$

The probability of which is:

-   In the susceptible state:

    $$
    P(y_t = \varnothing | Z_t = S) = 1 - \phi_S
    $$

-   In the infected state

    $$
    P(y_t = \varnothing | Z_t = I) = 1-\phi_I
    $$

## vii. Probability of a Ct value

Where a Ct value is observed, it can only happen if a detection has occurred, which means that the probability of seeing a Ct value has two components:

1.  probability that detection happened
2.  that the Ct took that value

The probability of which is:

-   In the susceptible state:

    $$
    p(y_t = c_t | Z_t = S) = \phi_S \mathcal{N}(C_t | \mu_S, \sigma_S)
    $$

-   In the infected state:

    $$
    p(y_t = c_t | Z_t = I) = \phi_I \mathcal{N}(C_t | \mu_I, \sigma_I)
    $$

$\mathcal{N}$ = the Gaussian distribution

$c_t$ = Ct value

$\mu_S, \mu_I$ = mean Ct values (S and I state)

$p$ is lowercase because this is a density not a probability mass (Ct is continuous and can be any number from 0 to infinity and anything in between, so it is a mass not a discrete probability, unlike PCR negative or positive!)

## viii. Emission model summary

-   Susceptible state:

    $$
    P(y_t \mid Z_t = S) =\begin{cases}1 - \phi_S, & y_t = \varnothing, \\\phi_S \, \mathcal{N}\!\left(y_t \mid \mu_S, \sigma_S\right), & y_t \in \mathbb{R}\end{cases}
    $$

In words (S state):

*"With probability* $1 - \phi_S$ *no amplification is detected; with probability* $\phi_S$*, amplification occurs, and the observed Ct value is drawn from a Normal distribution with mean* $\mu_S$ *and standard deviation* $\sigma_S$*."*

-   Infected state:

    $$
    P(y_t \mid Z_t = I) =\begin{cases}1 - \phi_I, & y_t = \varnothing, \\\phi_I \, \mathcal{N}\!\left(y_t \mid \mu_I, \sigma_I\right), & y_t \in \mathbb{R}\end{cases}
    $$

    In words (I state)

    With probability $1-\phi_I$, no amplification is detected; with probability $\phi_I​$, amplification occurs and the observed Ct value is drawn from a Normal distribution with mean $\mu_I$ and standard deviation $\sigma_1$

    ## ix. Time index and latent state process.

    $i = 1,...,N_{subjects}$ (1:336 index participants)

    $t = 1,...,T_i$ (index days within each individual $i$)

    [Initial state distribution]{.underline}

    $$
    P(Z_{i,1} = I) = \pi_I,\quad P(Z_{i,1} = S) = 1- \pi_I.
    $$

    where, $\pi_I$ is the probability of the infected state (here at the start of follow-up). So in words.

    *"The probability that subject* $i$ is in the infected state at $t=1$ is $\pi_I$"

    *"The probability of subject* $i$ *being susceptible at* $t=1$ is $1-\pi_I"$

    Even more simply, before I observe any Ct values, I still have to say what fraction of people I believe to be infected at the start of follow-up. Where $\pi_I$ is large, this is many people; where it is small, it is few people.

## x. SIS transition model

Let:

$$
p_{SI} = P(Z_{i,t+1} | Z_{i,t} = S), \quad p_{IS} = P(Z_{i,t,+1} = S | Z_{i,t} = I).
$$

*"The transition probability* $p_{SI}$ *is defined as the probability that subject* $i$ *is in the infected state at time* $t+1$*, given that they were in the susceptible state at time* $t$*"*

and

*"The transition probability* $p_{IS}$ *is defined as the probability that subject* $i$ *is in the susceptible state at time* $t+1$*, given that they were in the infected state at time* $t$*"*

Then for $t-1,..., T_i - 1$:

$$
P(Z_{i,t+1} = I | Z_{i,t} = S) = p_{SI}, \quad P(Z_{i,t+1} = S | Z_{i,t} = S) = 1 - p_{SI},\\ P(Z_{i,t+1} = S | Z_{i,t} = I) = p_{IS}, \quad P(Z_{i,t+1} = I | Z_{i,t} = I) = 1 - p_{IS}.
$$

And written as a transition matrix in state order $S,I$:

$$
\mathbf{T} =\begin{pmatrix}1 - p_{SI} & p_{SI} \\p_{IS} & 1 - p_{IS}\end{pmatrix}.
$$

## xi. Full joint likelihood

Now we join the state transition probabilities with the emission probabilities.

-   $R_{i,t}$ : whether a PCR test was performed at time $t$ (remembering that there are missing samples).

-   $D_{i,t}$: Ct value detected $D_{i,t} = 1$ or not detected $D_{i,t} = 0$

-   $O_i$ : is the set of observed (sampled) times for individual $i$

So the observation indicator is:

$$
R_{i,t} =\left\{\begin{array}{ll}1 & \text{if sampled at time } t \\0 & \text{otherwise}\end{array}\right.
$$

The set of observed days is:

$$
O_i = \{t: R_{i,t} = 1\}
$$

*(The set of days on which individual* $i$ *actually gave a sample)*

Observed data $y_{it}$ which may be either a Ct value ($\in \mathbb{R}$) or no amplification ($\varnothing$)

Therefore the individual level joint probability is:

$$
p(Z_{i,1:T_i},y_i, O_i)
$$

*Which means: "The probability that individual* $i$ *had a particular entire infection history AND produced the observations we saw on the days that we sampled them*

*This is the [likelihood contribution of one person.]{.underline}*

### *The full expression*

### 

$$
p(Z_{i,1:T_i},y_i,O_i) = p(Z_{i,1}) \prod*{t=1}^{{T_1} -1} p(Z*{i,t+1} | Z_{i,t}) \prod_{t \in O_i} p(y{i,t} | Z_{i,t})
$$

### Breaking it down:

#### A. Initial state

-   the probability that individual $i$ starts in state $S$ or $I$:

    $$
    P(Z_{i,1} = I) = \pi_I
    $$

(this is where we derive $\pi_i$ from)

#### B. State evolution over time

-   Multiply together the probabilities of ever day-to-day transition in the person's latent infection history

    $$
    \prod_{t=1}^{T_{i} - 1} p(Z_{i,t,+1} | Z_{i,t})
    $$

For example, let's say latent infection history had been: S-S; S:I; I:I; I:S... each step uses $p_{SI}, p_{I,S}$

The product of this builds the probability of the entire latent trajectory.

#### C. Observation model (only when sampled, though!)

$$
\prod_{t \in O_i} p(y_{i,t} | Z_{i,t})
$$

This says:

*"if the person was in state* $Z_{i,t}$, how likely is the observed Ct (or non-detection)?

#### D. Putting it together

for one person, this model says:

1.  They start in either $S$ or $I$
2.  Their state evolves day by day via the SIS process
3.  On days where I sampled them, their latent state generates the observed data (Ct or no amplification)

We assume here that individuals are conditionally independent, therefore the total likelihood is the product of each person's likelihood.

#### E. Across individuals

-   Here we assume conditional independence:

    $$
    p(Z,y) = \prod_{i=1}^{N_{subj}} p(Z_{i,1:T{_i}}, y_i, O_i).
    $$

### xii. Parameters (on the logit scale)

In the model, some of the parameters are probabilities. E.g. probability of recovering today...

The rule of probabilities is:

$$
0 ≤ p ≤ 1
$$

Probabilities can never be negative and never be bigger than one.

When the model is exploring parameter space it does so by proposing new values and moving freely in all directions, but the probability is in "a box with walls":

0 ├────────────── p ──────────────┤ 1

This means that the sampler bumps into boundaries and sampling is inefficient.

The key concept is that instead of working on the 0:1 space, we use the logit, which represents it

$$
\alpha = \log\!\left(\frac{p}{1 - p}\right)
$$

The logit stretches 0 and 1 out to the whole number line:

| probability $p$ | Logit $\alpha$ |
|-----------------|----------------|
| 0.01            | -4.6           |
| 0.1             | -2.2           |
| 0.5             | 0              |
| 0.9             | +2.2           |
| 0.99            | +4.6           |

Mathematically:

$$
\alpha \in (-\infty, + \infty)
$$

So now there are now boundaries in what stan can sample from.

### xiii. Expressing probabilities as logits.

Since we are using the logit scale, we need to rewrite our parameters

$$
logit(p_{SI}) = \alpha_{SI}, \quad logit(p_{IS}) = \alpha_{IS}, \quad logit(\pi_I) = \alpha_{\pi}, \\ logit(\phi_I) = \alpha_{\phi I}, \quad logit(\phi_S) = \alpha_{\phi_S}.
$$

### xiv. Defining our priors

Table \[x\]. Prior specification and interpretation for the Bayesian SIS hidden Markov model. Transition, initial state, and detection parameters are modelled on the logit scale; Ct distribution parameters are modelled on the natural Ct scale.

| Stan parameter | Scale prior is placed on | Implied model quantity | Prior specification | Biological interpretation |
|---------------|---------------|---------------|---------------|---------------|
| α_SI | Logit scale | p_SI = logit⁻¹(α_SI) | α_SI \~ Normal(logit(0.01), 0.7) | Daily infection hazard: probability of transitioning from susceptible to infected between consecutive days. Larger values imply more frequent new infections. |
| α_IS | Logit scale | p_IS = logit⁻¹(α_IS) | α_IS \~ Normal(logit(0.17), 0.5) | Daily recovery hazard: probability of clearing infection. Controls expected infection duration (approximately 1 / p_IS days). |
| α_π | Logit scale | π_I = logit⁻¹(α_π) | α_π \~ Normal(logit(0.063), 0.4) | Baseline prevalence: probability an individual is infected at the first observed day of follow-up. |
| α_ϕI | Logit scale | ϕ_I = logit⁻¹(α_ϕI) | α_ϕI \~ Normal(logit(0.75), 0.35) | PCR sensitivity: probability of observing a Ct value on a sampled day given true infection. |
| α_ϕS | Logit scale | ϕ_S = logit⁻¹(α_ϕS) | α_ϕS \~ Normal(logit(0.005), 0.8) | False positive probability: probability of observing a Ct value on a sampled day when uninfected (e.g. contamination or late-cycle amplification). |
| μ_I | Natural Ct scale | Used directly | μ_I \~ Normal(28, 2), constrained to \[10, 45\] | Mean Ct value in the infected state. Lower values correspond to higher pathogen load and earlier amplification. |
| μ_S | Natural Ct scale | Used directly | μ_S \~ Normal(40, 2), constrained to \[μ_I + 0.5, 45\] | Mean Ct value in the susceptible state, representing background or spurious amplification. Constraint enforces separation from infected Ct values. |
| σ_I | Natural Ct scale (positive) | Used directly | σ_I \~ Normal(4, 0.5), truncated to \[1, 6\] | Variability of Ct values in the infected state. Larger values imply greater heterogeneity in pathogen load. |
| σ_S | Natural Ct scale (positive) | Used directly | σ_S \~ Normal(3, 0.5), truncated to \[1, 6\] | Variability of background Ct values when false detection occurs. Controls dispersion of noise Ct values. |

### Priors (in prose!)

$α_{SI}​∼N(logit(0.01),0.7)$

This prior says that new infections are uncommon on any given day (about 1% per day on average), but I’m quite unsure about the exact rate and I’m happy for the data to move it by a few-fold in either direction.

$α{IS}∼N(logit(0.17),0.5)$

This says that once someone is infected they usually clear infection over several days rather than immediately, with a typical daily recovery chance around 17%, but with room for shorter or longer infections if the data support it.

$α_π​∼N(logit(0.063),0.4)$

This encodes a belief that only a small minority of people are already infected at the start of follow-up (around 6%), while allowing some uncertainty about the true baseline prevalence.

$α_{ϕ_I}​∼N(logit(0.75),0.35)$

This reflects the idea that PCR usually detects infection when someone is truly infected, but not perfectly, and allows for a reasonable amount of uncertainty in sensitivity across days and individuals.

$α_{ϕ_S}​∼N(logit(0.005),0.8)$

This says that false positive Ct values should be very rare when someone is uninfected (around half a percent), but I’m deliberately uncertain by a factor of a few because contamination and late-cycle noise do occasionally happen.

$μ_I​∼N(28,2)$ constrained to \[10, 45\]

This says that true infections tend to produce mid-range Ct values centred around 28, while ruling out implausibly low or high Ct means that would not make biological or laboratory sense.

$μ_S​∼N(40,2)$, constrained to $μ_S ​≥ μ_I​+0.5$

This encodes that background or spurious amplification, when it occurs, tends to appear very late in the PCR run, and I explicitly force it to sit above the infected Ct regime so the two states can’t swap roles.

$σ_I​∼N(, 4,0.5)$ ,truncated to \[1, 6\]

This allows moderate variability in Ct values among infected observations, reflecting heterogeneity in bacterial load, but prevents the model inventing unrealistically tight or wildly diffuse infected Ct distributions.

$σ_S​∼N(3,0.5)$ truncated to \[1, 6\]

This assumes that false positive Ct values cluster fairly tightly around a high Ct mean, while still allowing some spread, and stops the noise distribution becoming pathological.

## 1. Libraries

```{r}
library(tidyverse)
library(here)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(patchwork)
```

## 2. Read in modelling dataframe (not Ct value filtering)

```{r}
pcr <- read_csv(here::here("data_clean", "HMM_ModelA_input_dailygrid.csv")) %>%
  mutate(
    has_obs = as.integer(has_obs),
    has_ct  = as.integer(has_ct),
    Ct      = if_else(has_ct == 1L, ct_value, 0)  # placeholder when no Ct
  ) %>%
  arrange(pid_id, collection_date, row)
```

## 3. Make ID table, define Ct max and list stan data

```{r}


idx_tbl_A <- pcr %>%
  group_by(pid_id) %>%
  summarise(
    start = min(row),
    end   = max(row),
    .groups = "drop"
  )

Ct_max <- 45  # fixed detection limit

stan_data_A <- list(
  N         = nrow(pcr),
  N_subj    = nrow(idx_tbl_A),
  start_idx = idx_tbl_A$start,
  end_idx   = idx_tbl_A$end,
  has_obs   = as.array(pcr$has_obs),
  has_ct    = as.array(pcr$has_ct),
  Ct        = as.array(pcr$Ct),
  Ct_max    = Ct_max
)


```

## 4. Stan model

```{r}
stan_file_A <- here::here(
  "bayesian_models", "stan", "model_A_individual_hmm.stan"
)
stopifnot(file.exists(stan_file_A))

```

## 5. Fit model A and extract parameter summaries

```{r}

logit <- function(p) log(p / (1 - p))

init_fun <- function() {
  list(
    alpha_SI     = logit(0.01),
    alpha_IS     = logit(0.17),
    alpha_pi     = logit(0.063),

    alpha_phi_I  = logit(0.75),
    alpha_phi_S  = logit(0.005),

    mu_I         = 28,
    mu_S         = 40,
    sigma_I      = 4,
    sigma_S      = 3
  )
}

mod_A <- cmdstan_model(stan_file_A, force_recompile = TRUE)

fit_A <- mod_A$sample(
  data            = stan_data_A,
  init            = init_fun,
  chains          = 4,
  parallel_chains = 4,
  iter_warmup     = 2000,
  iter_sampling   = 2000,
  seed            = 2027,
  refresh         = 200,
  adapt_delta     = 0.95,
  max_treedepth   = 12
)



# Summarise key parameters (using names that exist in Stan)


pars_summary <- c(
  # logits 
  "alpha_SI", "alpha_IS", "alpha_pi",
  "alpha_phi_I", "alpha_phi_S",

  # probabilities on natural scale (transformed parameters)
  "p_SI", "p_IS", "pi_I", "phi_I", "phi_S",

  # Ct distribution parameters
  "mu_I", "mu_S", "sigma_I", "sigma_S"
)

results_model_A <- fit_A$summary(pars_summary) %>%
  dplyr::select(variable, mean, median, sd, q5, q95, rhat, ess_bulk, ess_tail)

results_model_A



```

## 6. Extract posterior mean

```{r}

# Extract posterior draws and summarise to the mean per day
prob_I_A <- fit_A$draws("prob_I", format = "draws_df") %>%
  summarise(across(everything(), mean)) %>%
  pivot_longer(
    cols      = everything(),
    names_to  = "var",
    values_to = "prob_I"
  ) %>%
  mutate(
    row = as.integer(gsub("prob_I\\[|\\]", "", var))
  ) %>%
  dplyr::select(row, prob_I)

```

## 7. Traceplots

```{r}

# Key parameters for traceplots
pars_core <- c("p_SI", "p_IS", "pi_I", "mu_I", "mu_S", "sigma_I", "sigma_S")

# Extract draws in matrix form (iterations x chains for each parameter)
draws_core <- fit_A$draws(pars_core)

# 1) Quick traceplots
bayesplot::mcmc_trace(draws_core, pars = pars_core)

#  separate panels for clarity:
bayesplot::mcmc_trace(draws_core, pars = c("p_SI", "p_IS", "pi_I"))
bayesplot::mcmc_trace(draws_core, pars = c("mu_I", "mu_S"))
bayesplot::mcmc_trace(draws_core, pars = c("sigma_I", "sigma_S"))

# 2) Rhat & ESS overview
fit_A$summary(pars_core)[, c("variable", "mean", "sd", "rhat", "ess_bulk", "ess_tail")]

```

## 8. WAIC

```{r}

# draws in matrix form: iterations*chains rows, N_subj columns
log_lik_mat <- fit_A$draws("log_lik_subj", format = "matrix")

dim(log_lik_mat)
# [1] n_draws   N_subj
waic_A <- loo::waic(log_lik_mat)

waic_A
print(waic_A)

# summary table
waic_summary <- with(waic_A, 
  tibble::tibble(
    waic      = waic,
    se_waic   = se_waic,
    p_waic    = p_waic,
    elpd_waic = elpd_waic,
    se_elpd   = se_elpd_waic
  )
)
waic_summary

loo_A <- loo::loo(log_lik_mat)
loo_A
print(loo_A)


```

## 9. Model diagnostics

```{r}
# CmdStan sampler diagnostics summary
fit_A$cmdstan_diagnose()

# Inspect divergences / treedepth from draws
draws <- fit_A$draws(variables = NULL)
np <- fit_A$sampler_diagnostics()
div <- np[, , "divergent__"]
td  <- np[, , "treedepth__"]

cat("Divergences:", sum(div), "\n")
cat("Max treedepth hits:", sum(td >= 12), "\n")



```

## 10. Predictive plots

### 10a. For one participant

```{r}


# 1. Join pcr and posterior infection probabilities -----------------------

pcr_with_prob <- pcr %>%
  left_join(prob_I_A, by = "row") %>%
  mutate(
    ct_value_num = suppressWarnings(as.numeric(ct_value)),
    pcr_pos = if_else(
      has_ct == 1L & !is.na(ct_value_num) & ct_value_num <= 37.1,
      1,  # PCR positive
      0   # PCR negative, no Ct, or Ct > 37.1
    )
  )

# 2. Subset to participant T27970 and sort by date ------------------------

predictive_plot_T27970 <- pcr_with_prob %>%
  filter(pid == "T27970") %>%      
  arrange(collection_date)

summary(predictive_plot_T27970$prob_I)

# 3. Plot: binary PCR positivity and prob_I -------------------------------

plot_T27970 <- ggplot(predictive_plot_T27970, aes(x = collection_date)) +
  
  # Binary PCR positivity (1 = Ct <= 37.1, 0 = otherwise)
  geom_point(
    aes(y = pcr_pos),
    colour = "firebrick",
    size   = 3
  ) +
  
  # Latent infection probability
  geom_line(
    aes(y = prob_I),
    colour    = "steelblue",
    linewidth = 1.1
  ) +
  
  labs(
    title = "Participant T27970 — PCR positivity (0/1) and inferred infection probability",
    x     = "Collection date",
    y     = "PCR positive (≤37.1) = 1   |   Negative / no Ct = 0"
  ) +
  
  theme_bw() +
  theme(
    plot.title  = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

plot_T27970



```

### 10b. For ten participants (PCR cut off ≤ 37.1)

In this chunk we define the PCR positives by our previously defined "cut of" of a Ct value of ≤ 37.1

```{r}

# 1. Join PCR dataframe with posterior probabilities


pcr_with_prob <- pcr %>%
  left_join(prob_I_A, by = "row") %>%
  mutate(
    ct_value_num = suppressWarnings(as.numeric(ct_value)),
    pcr_pos = if_else(
      has_ct == 1L & !is.na(ct_value_num) & ct_value_num <= 37.1,
      1,  # PCR positive
      0   # negative / no Ct
    )
  )


# 2. We choose ten participants with at least one PCR positive to visualise


ten_pids <- pcr_with_prob %>%
  group_by(pid) %>%
  summarise(
    n_pos = sum(pcr_pos == 1),
    n_obs = sum(has_obs == 1),
    .groups = "drop"
  ) %>%
  filter(n_pos > 0) %>%
  arrange(desc(n_pos), desc(n_obs)) %>%
  slice_head(n = 10) %>%
  pull(pid)

ten_pids



# 3. Define plotting function


plot_predictive <- function(pid, data) {

  df <- data %>%
    filter(pid == !!pid) %>%
    arrange(collection_date)

  ggplot(df, aes(x = collection_date)) +
    
    # PCR positivity (0/1)
    geom_point(
      aes(y = pcr_pos),
      size = 2.8,
      colour = "firebrick"
    ) +

    # Latent infection probability
    geom_line(
      aes(y = prob_I),
      linewidth = 1,
      colour = "steelblue"
    ) +

    labs(
      title = paste("Participant", pid),
      x = "Date",
      y = "PCR positive (Ct ≤ 37.1) = 1   |   Negative/no Ct = 0   +   P(Infected)"
    ) +

    theme_bw() +
    theme(
      plot.title  = element_text(size = 11, face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    
    coord_cartesian(ylim = c(0, 1))   # make everything visible on the same scale
}


# 4. Produce 10 plots and vertically stack them


plot_list <- lapply(ten_pids, plot_predictive, data = pcr_with_prob)

combined_plot_10 <- wrap_plots(plotlist = plot_list, ncol = 1)

combined_plot_10

#  save
ggsave(
  here::here("outputs", "figures", "predictive_10_participants.tiff"),
  combined_plot_10,
  device = "tiff",
  dpi    = 600,
  width  = 10,
  height = 20,
  bg     = "white"
)

```

### 10c. Combined predictive plot with Ct values \> 37.1 displayed seperately

```{r}


# 1. Define updated plotting function where Ct value > 37.1 displayed separately


plot_predictive_2 <- function(pid, data) {

  df <- data %>%
    filter(pid == !!pid) %>%
    arrange(collection_date)

  ggplot(df, aes(x = collection_date)) +

    # PCR positive (Ct ≤ 37.1)
    geom_point(
      data = df %>% filter(has_ct == 1L, !is.na(ct_value_num), ct_value_num <= 37.1),
      aes(y = 1),
      size = 2.8,
      colour = "firebrick"
    ) +

    # PCR amplification detected but Ct > 37.1
    geom_point(
      data = df %>% filter(has_ct == 1L, !is.na(ct_value_num), ct_value_num > 37.1),
      aes(y = 1),
      size = 2.8,
      colour = "orange"
    ) +

    # Latent infection probability
    geom_line(
      aes(y = prob_I),
      linewidth = 1,
      colour = "steelblue"
    ) +

    labs(
      title = paste("Participant", pid),
      x = "Date",
      y = "PCR: Ct ≤ 37.1 (red), Ct > 37.1 (orange)   +   P(Infected)"
    ) +

    theme_bw() +
    theme(
      plot.title  = element_text(size = 11, face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    coord_cartesian(ylim = c(0, 1))
}


# 2. Produce and stack plots


plot_list <- lapply(ten_pids, plot_predictive_2, data = pcr_with_prob)

combined_plot_10 <- wrap_plots(plotlist = plot_list, ncol = 1)

combined_plot_10


# 3. Save


ggsave(
  here::here("outputs", "figures", "predictive_10_participants_ct_thresholds.tiff"),
  combined_plot_10,
  device = "tiff",
  dpi    = 600,
  width  = 10,
  height = 20,
  bg     = "white"
)

```

## 11. Optimising priors for subsequent models

```{r}
# =========================================================
# Model A → Prior-tuning diagnostics (end-to-end)
# Prints the summaries we need to inform Model E priors
# Assumes you already have: pcr_with_prob (as shown), with:
# pid, collection_date, row, has_obs, has_ct, ct_value_num, pcr_pos, prob_I
# =========================================================

library(dplyr)
library(lubridate)
library(data.table)

# --- 0) Quick structure sanity checks ------------------------------------
cat("\n==================== 0) Sanity checks ====================\n")

required_cols <- c(
  "pid","collection_date","row","has_obs","has_ct",
  "ct_value_num","pcr_pos","prob_I"
)

missing_cols <- setdiff(required_cols, names(pcr_with_prob))
if (length(missing_cols) > 0) {
  stop("pcr_with_prob is missing columns: ", paste(missing_cols, collapse = ", "))
}

cat("Rows:", nrow(pcr_with_prob), "\n")
cat("Unique participants (pid):", dplyr::n_distinct(pcr_with_prob$pid), "\n")

# --- 1) Sampling cadence and follow-up length ----------------------------
cat("\n==================== 1) Sampling cadence by pid ====================\n")

cadence_by_pid <- pcr_with_prob %>%
  group_by(pid) %>%
  summarise(
    date_min = min(collection_date, na.rm = TRUE),
    date_max = max(collection_date, na.rm = TRUE),
    n_days_grid = n(),
    n_obs = sum(has_obs == 1),
    obs_rate = n_obs / n_days_grid,
    median_gap_days_between_obs = {
      d <- sort(unique(collection_date[has_obs == 1]))
      if (length(d) < 2) NA_real_ else median(as.numeric(diff(d)), na.rm = TRUE)
    },
    .groups = "drop"
  )

print(summary(cadence_by_pid$obs_rate))
cat("\nMedian gap (days) between observed tests (per pid):\n")
print(summary(cadence_by_pid$median_gap_days_between_obs))

cat("\nTop 10 pids by number of observed tests:\n")
print(cadence_by_pid %>% arrange(desc(n_obs)) %>% slice_head(n = 10))

# --- 2) Positivity rates on TESTED days only -----------------------------
cat("\n==================== 2) Positivity on tested days (has_obs==1) ====================\n")

positivity_overall <- pcr_with_prob %>%
  filter(has_obs == 1) %>%
  summarise(
    n_tests = n(),
    any_ct_rate = mean(has_ct == 1, na.rm = TRUE),
    ct_le_37_rate = mean(pcr_pos == 1, na.rm = TRUE),
    .groups = "drop"
  )

print(positivity_overall)

cat("\nPositivity by pid (distribution):\n")
positivity_by_pid <- pcr_with_prob %>%
  filter(has_obs == 1) %>%
  group_by(pid) %>%
  summarise(
    n_tests = n(),
    any_ct_rate = mean(has_ct == 1),
    ct_le_37_rate = mean(pcr_pos == 1),
    .groups = "drop"
  )

print(summary(positivity_by_pid$any_ct_rate))
print(summary(positivity_by_pid$ct_le_37_rate))

# --- 3) Ct distribution among detected tests -----------------------------
cat("\n==================== 3) Ct distribution among detected tests (has_obs==1 & has_ct==1) ====================\n")

ct_dist <- pcr_with_prob %>%
  filter(has_obs == 1, has_ct == 1) %>%
  summarise(
    n_ct = n(),
    min = min(ct_value_num, na.rm = TRUE),
    q25 = quantile(ct_value_num, 0.25, na.rm = TRUE),
    median = median(ct_value_num, na.rm = TRUE),
    mean = mean(ct_value_num, na.rm = TRUE),
    q75 = quantile(ct_value_num, 0.75, na.rm = TRUE),
    max = max(ct_value_num, na.rm = TRUE),
    .groups = "drop"
  )

print(ct_dist)

cat("\nCt bins (detected tests only):\n")
ct_bins <- pcr_with_prob %>%
  filter(has_obs == 1, has_ct == 1) %>%
  mutate(
    ct_bin = case_when(
      ct_value_num <= 30 ~ "<=30",
      ct_value_num <= 37.1 ~ "30–37.1",
      ct_value_num <= 40 ~ "37.1–40",
      TRUE ~ "40–45"
    )
  ) %>%
  count(ct_bin) %>%
  mutate(prop = n / sum(n)) %>%
  arrange(match(ct_bin, c("<=30","30–37.1","37.1–40","40–45")))

print(ct_bins)

# --- 4) High Ct isolation vs clustering on TESTED days -------------------
cat("\n==================== 4) High Ct runs on tested days (Ct > 37.1) ====================\n")

dt <- as.data.table(pcr_with_prob)
setorder(dt, pid, row)

dt[, tested := (has_obs == 1)]
dt[, any_ct := tested & (has_ct == 1)]
dt[, high_ct := tested & (has_ct == 1) & !is.na(ct_value_num) & (ct_value_num > 37.1)]
dt[, low_ct  := tested & (has_ct == 1) & !is.na(ct_value_num) & (ct_value_num <= 37.1)]
dt[, no_ct   := tested & (has_ct == 0)]

dt_tested <- dt[tested == TRUE,
                .(pid, row, collection_date, ct_value_num, high_ct, low_ct, any_ct, no_ct, prob_I)]
setorder(dt_tested, pid, row)

# run-length encoding on tested days
dt_tested[, run_id_high := rleid(high_ct), by = pid]
high_runs <- dt_tested[high_ct == TRUE, .N, by = .(pid, run_id_high)]
setnames(high_runs, "N", "run_len_tested_days")

cat("\nSummary of high-Ct run length (in tested days):\n")
print(summary(high_runs$run_len_tested_days))

cat("\nHigh-Ct run length table:\n")
print(table(high_runs$run_len_tested_days))

high_run_summary <- high_runs[, .(
  n_runs = .N,
  prop_isolated = mean(run_len_tested_days == 1),
  prop_clustered_ge2 = mean(run_len_tested_days >= 2)
)]
cat("\nHigh-Ct run summary:\n")
print(high_run_summary)

# --- 5) Proxy phi_I / phi_S using posterior prob_I thresholds ------------
cat("\n==================== 5) Proxy detection rates using prob_I thresholds ====================\n")

thr_I <- 0.8
thr_S <- 0.2

dt_tested[, infected_proxy := (prob_I >= thr_I)]
dt_tested[, susceptible_proxy := (prob_I <= thr_S)]

phi_I_proxy <- dt_tested[infected_proxy == TRUE, mean(any_ct)]
phi_S_proxy <- dt_tested[susceptible_proxy == TRUE, mean(any_ct)]

cat("Thresholds: infected_proxy prob_I >= ", thr_I, " ; susceptible_proxy prob_I <= ", thr_S, "\n", sep = "")
cat("Proxy phi_I (P(any Ct | prob_I high)):", phi_I_proxy, "\n")
cat("Proxy phi_S (P(any Ct | prob_I low )):", phi_S_proxy, "\n")

cat("\nCt means conditional on proxy state (detected only):\n")
muI_proxy <- dt_tested[infected_proxy == TRUE & any_ct == TRUE, mean(ct_value_num, na.rm = TRUE)]
muS_proxy <- dt_tested[susceptible_proxy == TRUE & any_ct == TRUE, mean(ct_value_num, na.rm = TRUE)]
cat("Proxy mu_I (detected, prob_I high):", muI_proxy, "\n")
cat("Proxy mu_S (detected, prob_I low ):", muS_proxy, "\n")

# --- 6) Episode-style summaries on DAILY grid using prob_I threshold -----
cat("\n==================== 6) Proxy infected run lengths + episode start rate (daily grid) ====================\n")

thr_run <- 0.5
dt2 <- copy(dt)
setorder(dt2, pid, row)

dt2[, I_proxy := (prob_I >= thr_run)]
dt2[, I_prev := shift(I_proxy, 1, type = "lag"), by = pid]

# Run lengths (in daily rows)
dt2[, I_run := rleid(I_proxy), by = pid]
I_runs <- dt2[I_proxy == TRUE, .N, by = .(pid, I_run)]
setnames(I_runs, "N", "run_len_days")

cat("\nProxy infected run length (days) summary (thr_run = ", thr_run, "):\n", sep = "")
print(summary(I_runs$run_len_days))

cat("\nProxy infected run length table (days):\n")
print(table(I_runs$run_len_days))

# Episode start rate per S-day (daily grid)
starts <- dt2[!is.na(I_prev) & I_prev == FALSE & I_proxy == TRUE, .N]$N
S_days <- dt2[!is.na(I_prev) & I_prev == FALSE, .N]$N
episode_start_rate <- starts / S_days

cat("\nProxy episode start rate per susceptible day (daily grid):\n")
cat("starts =", starts, " ; S_days =", S_days, " ; rate =", episode_start_rate, "\n")

# --- 7) Short printed takeaways to guide prior choices -------------------
cat("\n==================== 7) Takeaways to guide Model E priors ====================\n")
cat("* any_ct_rate on tested days informs phi_I/phi_S mixture and overall detectability.\n")
cat("* high-Ct run isolation vs clustering informs whether late Ct is mostly noise (phi_S up) or persistent I (p_IS down).\n")
cat("* proxy run lengths (daily) inform plausible p_IS (duration ~ 1/p_IS).\n")
cat("* proxy episode_start_rate informs plausible p_SI range (but not equal to any_ct_rate).\n")

```

## 12. Save everything here to pick up in script 18

In script 18 I will change the priors to more closely align with the outputs of chunk \[11\] above.

This chunk will save everything I need to quickly make "Model E" (which will be exactly as model A is, but with these modified priors)

```{r}


# --- directories ----------------------------------------------------------
rds_dir <- here::here("outputs", "rds")
fit_dir <- here::here("outputs", "fits")

dir.create(rds_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(fit_dir, showWarnings = FALSE, recursive = TRUE)

# --- 12a) Save core state (lightweight; no big fit objects inside) --------
state_17 <- list(
  # core data layer
  pcr         = pcr,
  idx_tbl_A   = idx_tbl_A,
  stan_data_A = stan_data_A,

  # model text and file path (so Script 18 can reproduce Model A exactly)
  stan_file_A       = stan_file_A,
  stan_model_A_text = stan_model_A_text,

  # key derived objects (useful for plotting / diagnostics without refitting)
  prob_I_A      = prob_I_A,
  pcr_with_prob = pcr_with_prob,

  # diagnostics from block 11 (if present)
  cadence_by_pid    = if (exists("cadence_by_pid")) cadence_by_pid else NULL,
  positivity_by_pid = if (exists("positivity_by_pid")) positivity_by_pid else NULL,

  # file paths for heavy objects saved separately
  fit_A_path       = here::here("outputs", "fits", "fit_A.rds"),
  log_lik_mat_path = here::here("outputs", "rds", "log_lik_mat_D.rds"),
  loo_A_path        = here::here("outputs", "rds", "loo_A.rds"),
  waic_A_path       = here::here("outputs", "rds", "waic_A.rds")
)

saveRDS(state_17, file = file.path(rds_dir, "model_A_state_core.rds"))
cat("Saved core state to:", file.path(rds_dir, "model_A_state_core.rds"), "\n")

# --- 12b) Save Model A fit object (for direct comparison in Script 18) ----
if (exists("fit_A")) {
  saveRDS(fit_A, file = file.path(fit_dir, "fit_A.rds"))
  cat("Saved fit_A to:", file.path(fit_dir, "fit_A.rds"), "\n")
} else {
  cat("NOTE: fit_A not found in environment; did not save fit_A.rds\n")
}

# --- 12c) Save log-likelihood matrix used for LOO/WAIC --------------------
# Prefer using existing log_lik_mat if you already created it in chunk 8.
# Otherwise, build it from fit_A.
if (exists("log_lik_mat")) {
  log_lik_mat_D <- log_lik_mat
} else if (exists("fit_A")) {
  log_lik_mat_D <- fit_A$draws("log_lik_subj", format = "matrix")
} else {
  log_lik_mat_D <- NULL
}

if (!is.null(log_lik_mat_D)) {
  saveRDS(log_lik_mat_D, file = file.path(rds_dir, "log_lik_mat_D.rds"))
  cat("Saved log_lik_mat_D to:", file.path(rds_dir, "log_lik_mat_D.rds"), "\n")
} else {
  cat("NOTE: Could not save log_lik_mat_D (need log_lik_mat or fit_A)\n")
}

# --- 12d) Save LOO + WAIC objects (optional but convenient) ---------------
# If already computed, save them; otherwise compute from log_lik_mat_D.
if (!exists("loo_A") && !is.null(log_lik_mat_D)) {
  loo_A <- loo::loo(log_lik_mat_D)
}
if (!exists("waic_A") && !is.null(log_lik_mat_D)) {
  waic_A <- loo::waic(log_lik_mat_D)
}

if (exists("loo_A")) {
  saveRDS(loo_A, file = file.path(rds_dir, "loo_A.rds"))
  cat("Saved loo_A to:", file.path(rds_dir, "loo_A.rds"), "\n")
} else {
  cat("NOTE: loo_A not available; did not save loo_A.rds\n")
}

if (exists("waic_A")) {
  saveRDS(waic_A, file = file.path(rds_dir, "waic_A.rds"))
  cat("Saved waic_A to:", file.path(rds_dir, "waic_A.rds"), "\n")
} else {
  cat("NOTE: waic_A not available; did not save waic_A.rds\n")
}

```

## 95% credible interbals

```{r}
# -------------------------------------------------------------------------
# Summarise key parameters with 95% credible intervals
# -------------------------------------------------------------------------

pars_summary <- c(
  # logits 
  "alpha_SI", "alpha_IS", "alpha_pi",
  "alpha_phi_I", "alpha_phi_S",

  # probabilities on natural scale
  "p_SI", "p_IS", "pi_I", "phi_I", "phi_S",

  # Ct distribution parameters
  "mu_I", "mu_S", "sigma_I", "sigma_S"
)

# Extract posterior draws
draws_A <- fit_A$draws(
  variables = pars_summary,
  format = "draws_df"
)

# Summarise posterior draws using 95% credible intervals
results_model_A_95 <- draws_A %>%
  as_tibble() %>%
  dplyr::select(all_of(pars_summary)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) %>%
  group_by(variable) %>%
  summarise(
    mean   = mean(value),
    median = median(value),
    sd     = sd(value),
    q2.5   = quantile(value, 0.025),
    q97.5  = quantile(value, 0.975),
    .groups = "drop"
  )

# Add MCMC diagnostics from CmdStan summary
diagnostics_A <- fit_A$summary(pars_summary) %>%
  dplyr::select(variable, rhat, ess_bulk, ess_tail)

results_model_A_95 <- results_model_A_95 %>%
  left_join(diagnostics_A, by = "variable") %>%
  mutate(
    variable = factor(variable, levels = pars_summary)
  ) %>%
  arrange(variable)

results_model_A_95

##
# -------------------------------------------------------------------------
# Derived quantities with 95% credible intervals
# -------------------------------------------------------------------------

derived_draws_A <- fit_A$draws(
  variables = c("p_SI", "p_IS", "phi_I", "phi_S"),
  format = "draws_df"
) %>%
  as_tibble() %>%
  mutate(
    acquisition_percent      = 100 * p_SI,
    clearance_percent        = 100 * p_IS,
    mean_duration_days       = 1 / p_IS,
    detection_in_I_percent   = 100 * phi_I,
    amplification_in_S_percent = 100 * phi_S,
    specificity_like_percent = 100 * (1 - phi_S)
  ) %>%
  dplyr::select(
    acquisition_percent,
    clearance_percent,
    mean_duration_days,
    detection_in_I_percent,
    amplification_in_S_percent,
    specificity_like_percent
  )

derived_results_A_95 <- derived_draws_A %>%
  pivot_longer(
    cols = everything(),
    names_to = "quantity",
    values_to = "value"
  ) %>%
  group_by(quantity) %>%
  summarise(
    mean   = mean(value),
    median = median(value),
    sd     = sd(value),
    q2.5   = quantile(value, 0.025),
    q97.5  = quantile(value, 0.975),
    .groups = "drop"
  )

derived_results_A_95
```
