---
title: "Bayesian Model B: household-coupled hidden Markov model"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

# Overview

Model B extends the susceptible–infected hidden Markov framework to jointly model members of the same household. Daily acquisition probability depends on the number of other latently infected household members at the preceding time step, baseline household environmental contamination, a baseline individual mobility proxy (days leaving the house in the preceding week), and a household-level random effect.

The household likelihood is marginalised over joint latent states using a forward algorithm implemented in Stan, with `reduce_sum` used for within-chain parallelisation. The single household with 14 participants is excluded because the joint state space grows as $2^M$ with household size.

The full fit is computationally intensive and may require multi-day runtime depending on hardware. The workflow documents data construction, model specification, priors, sampler settings, posterior summaries, and derived quantities used for the manuscript.

## Mathematical notation

### Latent state definition

Each individual $i$ in household $h$ at time $t$ occupies a latent state:

$$
Z_{i,h,t} \in \{S,I\},
$$

Where $S$ denotes susceptible and $I$ denotes infected

### Initial state distribution

At the start of follow-up, latent infection states are assumed to be independent across household members:$$
P(Z_{i,h,1} = I) = \pi, \quad logit(\pi) = \alpha_\pi.
$$

$\alpha_\pi$ captures the baseline probability of infection at study entry.

### State transition model (daily Markov process)

Recovery ( $I$ to $S$)

$$
P(Z_{i,h,t} = S | Z_{i,h,t-1} = I) = p_{IS}, \quad logit(p_{IS}) = \alpha_{IS}
$$

Acquistion ( $S$ to $I$ ) model B

Define:

-   $I_{h,t-1}$ : number of infected household members at time $t-1$

-   $X_h$ : household environmental contamination indicator

-   $D_i \in \{0,...,7\}$ : days leaving the house in the last week (measured at baseline)

-   $D_i^{\*} = \frac{D_i - \bar{D}}{7}$ : centred and scaled community exposure

    -   $D_i$ : the original value for person $i$ (0:7)

    -   $D_{i}^{\*}$ : the centred and scaled version of $D_i$

    -   $\bar{D}$ : the mean of $D$ across individuals

-   $u_h \sim \mathcal{N}(0, \sigma_u)$ : household level random effect.

Then:

$$
P(Z_{i,h,t} = I | Z_{i,h,t-1} = S) = p_{SI,i,h,t},
$$

with:

$$
logit(p_{SI,i,h,t}) = \alpha_{SI} + \beta_{hh} I{h,t-1} + \beta_{env}X_h + \beta_{comm}D_i^{\*} + u_h.
$$

Interpretation: acquisition risk is decomposed into within-household transmission pressure, shared environmental exposure, baseline community exposure propensity, and residual household heterogeneity.

### Observation model: PCR detection and Ct values

Let $Y_{i,h,t}$ denote the PCR observation for individual $i$ at time $t$

Detection component

conditional on latent state:

$$
\Pr(\text{Ct observed} \mid Z_{i,h,t} = I) = \phi_I,\qquad\Pr(\text{Ct observed} \mid Z_{i,h,t} = S) = \phi_S,
$$

with

$$
\text{logit}(\phi_I) = \alpha_{\phi I},\qquad\text{logit}(\phi_S) = \alpha_{\phi S}.
$$

Interpretation: infected and uninfected states have different probabilities of yielding an amplifiable Ct value

Ct emission:

$$
C_{i,h,t} \mid Z_{i,h,t} = I \sim \mathcal{N}(\mu_I, \sigma_I^2),\qquad C_{i,h,t} \mid Z_{i,h,t} = S \sim \mathcal{N}(\mu_S, \sigma_S^2),
$$

with

$\mu_I < \mu_S$

interpretation: Ct values are treated as continuous measurements whose distributions differ by latent infection state.

### Full likelihood (household level)

For household $h$, the likelihood marginalised over all latent state sequences is

$$
L_h=\sum_{\mathbf{Z}_h}\Pr(\mathbf{Z}_{h,1})\prod_{t=2}^{T} \Pr(\mathbf{Z}_{h,t} \mid \mathbf{Z}_{h,t-1})\prod_{t=1}^{T} \Pr(\mathbf{Y}_{h,t} \mid \mathbf{Z}_{h,t}) .
$$

Where

-   $L_h$ is the likelihood contribution from household $h$

-   $\mathbf{Z}_{h,t}$ is the joint latent state vector of all individuals in household $h$ at time $t$

-   $\mathbf{Y}_{h,t}$ is the vector of observed data (Ct / PCR outcomes) for all indiividuals in household $h$ at time $t$

-   $\sum_{\mathbf{Z}_h}$ means sum over all posible latent state sequences for household $h$

Interpretation: the model integrates over all possible household-level transmission histories using the forward algorithm

### Community-exposure extension

Model B includes introducing a baseline individual level proxy for community exposure into the S to I transition model, whilst retaining the same household level structure and continuous Ct emission model.

### Reminders

Due to concerns about the explosion in state space, we are still going to exclude the household with 14 participants from the model.

## 1. Libraries

```{r}
library(tidyverse)
library(lubridate)
library(here)
library(cmdstanr)
library(posterior)
library(patchwork)
library(loo)
```

## 2. Load prepared household-model input

```{r}
setup_file <- here("rds", "model_B_prepared_input.rds")
if (!file.exists(setup_file)) {
  stop(
    "Missing rds/model_B_prepared_input.rds. ",
    "This prepared household-model input object is provided with the controlled-access analytic data package."
  )
}
obj <- readRDS(setup_file)

pcr_hmm        <- obj$pcr_hmm
id_pid         <- obj$id_pid
id_hh          <- obj$id_hh
idx_tbl        <- obj$idx_tbl
stan_data_base <- obj$stan_data_base
stan_data_base_extended    <- obj$stan_data_base_extended

rm(obj)
```

## 3. Read baseline mobility covariate

```{r}
CRF_B <- read_csv(here("data_raw", "CRF_B_MCA.csv"))
#
CRF_B <-  CRF_B %>% 
  select(hhid, data_date, pid, dys_leav_day)
```

## 4. Join

Here we attach the days leaving the house variable to every transition point

```{r}
mobility_lookup <- CRF_B %>%
  select(pid, dys_leav_day) %>%
  distinct()

stopifnot(nrow(mobility_lookup) == n_distinct(mobility_lookup$pid))

pcr_hmm <- pcr_hmm %>%
  left_join(mobility_lookup, by = "pid")

```

## 5. Centre and scale

-   `D_star = 0` → typical mobility

-   `D_star = +1` → maximal mobility

-   `D_star = −1` → minimal mobility

```{r}

# Person-level mean across individuals (NOT weighted by number of rows per pid)
D_bar <- mobility_lookup %>%
  summarise(D_bar = mean(dys_leav_day, na.rm = TRUE)) %>%
  pull(D_bar)



# Centred and scaled by 7
pcr_hmm <- pcr_hmm %>%
  mutate(
    D_star = (dys_leav_day - D_bar) / 7
  )

# Check: D_star is constant within pid

check_const <- pcr_hmm %>%
  group_by(pid) %>%
  summarise(n_unique = n_distinct(D_star), .groups = "drop") %>%
  summarise(max_unique = max(n_unique)) %>%
  pull(max_unique)

stopifnot(check_const == 1)


```

## 6. Build Model B arrays and Stan data

This chunk drops the single largest household (M = 14), re-indexes households and members, and constructs the `has_obs`, `has_ct`, and `Ct` arrays plus the precomputed joint-state table used by the household HMM with `reduce_sum`. It then **reuses the already-computed `D_star` in `pcr_hmm`**, maps it onto a household-member matrix `D_star[h, m]`, and packages everything into `stan_data_B`.

```{r}
# ---- Chunk: Build Model B arrays and Stan data (reuse precomputed D_star) ----

stopifnot(exists("pcr_hmm"), exists("id_hh"), exists("id_pid"))
stopifnot("D_star" %in% names(pcr_hmm))  # we already created this earlier

# 0) Identify the one household with M=14 (distinct members)
hh_sizes <- pcr_hmm %>%
  distinct(hhid_id, pid_id) %>%
  count(hhid_id, name = "M")

hhid_drop <- hh_sizes %>% filter(M == 14) %>% pull(hhid_id)
stopifnot(length(hhid_drop) == 1)
message("Dropping household hhid_id = ", hhid_drop, " (M=14)")

# 1) Clean long df and drop duplicated env cols (if present), then drop the M=14 household
df <- pcr_hmm %>%
  filter(hhid_id != hhid_drop) %>%
  select(-matches("^environmental_baseline_positive\\.[xy]$")) %>%
  mutate(
    pid_id  = as.integer(pid_id),
    hhid_id = as.integer(hhid_id),
    has_obs = as.integer(has_obs),
    has_ct  = as.integer(has_ct),
    Ct      = as.numeric(Ct),
    D_star  = as.numeric(D_star),
    collection_date = as.Date(collection_date)
  ) %>%
  arrange(hhid_id, pid_id, collection_date)

stopifnot(!any(is.na(df$D_star)))  # you said no missing; D_star is baseline-replicated

# 2) Reindex households to dense 1..H_B for arrays / Stan
hh_map <- df %>%
  distinct(hhid_id, hhid) %>%
  arrange(hhid_id) %>%
  mutate(h = row_number()) %>%
  select(hhid_id, hhid, h)

H_B <- nrow(hh_map)

df <- df %>%
  left_join(hh_map %>% select(hhid_id, h), by = "hhid_id") %>%
  arrange(h, pid_id, collection_date)

stopifnot(!any(is.na(df$h)))

# 3) X_h vector aligned to h=1..H_B
X_h_B <- id_hh %>%
  filter(hhid_id != hhid_drop) %>%
  left_join(hh_map, by = "hhid_id") %>%
  arrange(h) %>%
  pull(environmental_baseline_positive) %>%
  as.integer()

stopifnot(length(X_h_B) == H_B)

# 4) Roster: member_id within each household (h), stable by pid_id
roster_B <- df %>%
  distinct(h, pid_id, pid) %>%
  arrange(h, pid_id) %>%
  group_by(h) %>%
  mutate(member_id = row_number()) %>%
  ungroup()

M_h_B <- roster_B %>%
  count(h, name = "M_h") %>%
  arrange(h) %>%
  pull(M_h)

M_max_B <- max(M_h_B)
message("After exclusion: H_B=", H_B, ", M_max_B=", M_max_B, " (max states = 2^M_max_B)")

# 4b) D_star matrix [H_B, M_max_B] aligned to (h, member_id)
D_star_mat <- matrix(0.0, nrow = H_B, ncol = M_max_B)

# Pull each member's baseline D_star once (it is constant within pid)
roster_mob <- roster_B %>%
  left_join(df %>% distinct(pid, D_star), by = "pid") %>%
  arrange(h, member_id)

stopifnot(!any(is.na(roster_mob$D_star)))

for (r in seq_len(nrow(roster_mob))) {
  hh <- roster_mob$h[r]
  mm <- roster_mob$member_id[r]
  D_star_mat[hh, mm] <- roster_mob$D_star[r]
}

# 5) Day index within household (t = 1..T_K), aligned to household start date
hh_day0_B <- df %>%
  group_by(h) %>%
  summarise(day0 = min(collection_date), .groups = "drop")

df <- df %>%
  left_join(hh_day0_B, by = "h") %>%
  mutate(day = as.integer(collection_date - day0) + 1L) %>%
  select(-day0)

T_K <- max(df$day)

# 6) Attach member_id
df <- df %>%
  left_join(roster_B %>% select(h, pid_id, member_id), by = c("h", "pid_id"))

stopifnot(!any(is.na(df$member_id)))
stopifnot(all(df$member_id >= 1 & df$member_id <= M_max_B))

# 7) Allocate arrays required by Stan code
has_obs_arr <- array(0L, dim = c(H_B, M_max_B, T_K))
has_ct_arr  <- array(0L, dim = c(H_B, M_max_B, T_K))
Ct_arr      <- array(0.0, dim = c(H_B, M_max_B, T_K))

# 7a) Ensure Ct has no NAs (store 0 when has_ct==0; Stan gates on has_ct)
df <- df %>%
  mutate(
    Ct = if_else(has_ct == 1L, as.numeric(Ct), 0.0)
  )

stopifnot(!any(is.na(df$Ct)))

idx <- cbind(df$h, df$member_id, df$day)
has_obs_arr[idx] <- df$has_obs
has_ct_arr[idx]  <- df$has_ct
Ct_arr[idx]      <- df$Ct

# 8) Precompute joint state table for maximum household size
S_h_B  <- as.integer(2L ^ M_h_B)
S_maxB <- as.integer(max(S_h_B))

z_state_mat <- matrix(0L, nrow = S_maxB, ncol = M_max_B)
for (s in 0:(S_maxB - 1L)) {
  z_state_mat[s + 1L, ] <- as.integer(intToBits(s))[1:M_max_B]  # LSB = member 1
}

z_state_arr <- array(as.integer(z_state_mat), dim = c(S_maxB, M_max_B))

# 9) Build Stan data list for Model B (reduce_sum + z_state + D_star[h,m])
stan_data_B <- list(
  H = H_B,
  M_max = M_max_B,
  T = T_K,

  S_max = S_maxB,
  S_h   = S_h_B,
  z_state = z_state_arr,

  M_h = as.integer(M_h_B),
  X_h = as.integer(X_h_B),

  has_obs = has_obs_arr,
  has_ct  = has_ct_arr,
  Ct      = Ct_arr,

  D_star  = D_star_mat,   # NEW for Model B

  grainsize = 2L  # tune 2–6
)

str(stan_data_B)

# 10) Save setup
dir.create(here("outputs", "rds"), recursive = TRUE, showWarnings = FALSE)

saveRDS(
  list(
    df_long_K    = df,
    roster_B     = roster_B,
    hh_map       = hh_map,
    hh_day0_B    = hh_day0_B,
    stan_data_B  = stan_data_B,
    H_B          = H_B,
    M_max_B      = M_max_B,
    T_K          = T_K,
    S_maxB       = S_maxB,
    hhid_drop    = hhid_drop
  ),
  here("outputs", "rds", "model_B_setup.rds")
)

message("Saved Model B setup: ", here("outputs", "rds", "model_B_setup.rds"))


```

## 7. Alignment check

```{r}
# ---- Alignment check: D_star_mat[h, m] matches df$D_star ----

check_D_star <- df %>%
  distinct(h, member_id, pid, D_star) %>%
  mutate(
    D_star_mat = mapply(
      function(hh, mm) D_star_mat[hh, mm],
      h, member_id
    )
  )

stopifnot(
  all(abs(check_D_star$D_star - check_D_star$D_star_mat) < 1e-12)
)

message("Alignment check passed: D_star_mat[h, m] matches df$D_star for all members.")

```

## 8. Write and save model B

```{r}
# ---- Chunk: Write + save Model B Stan code (FORWARD-FIXED + SLICE-AWARE reduce_sum indexing) ----

library(here)

stan_file_B <- here(
  "bayesian_models", "stan", "model_B_household_chmm.stan"
)
stopifnot(file.exists(stan_file_B))

```

## 9. Compile and fit model B

```{r}
# ---- Chunk: Compile + fit Model B (threaded sampling configuration) ----

library(cmdstanr)
library(here)

stopifnot(exists("stan_data_B"))

stan_file_B <- here("bayesian_models", "stan", "model_B_household_chmm.stan")
stopifnot(file.exists(stan_file_B))

# Threading configuration
threads <- 2              # safe, efficient
Sys.setenv(STAN_NUM_THREADS = threads)

stan_data_B$grainsize <- 1L   

mod_B <- cmdstan_model(
  stan_file_B,
  cpp_options = list(stan_threads = TRUE),
  force_recompile = TRUE
)

init_fun_B <- function() {
  list(
    alpha_SI  = qlogis(0.04),
    beta_env  = 0,
    beta_hh   = 0,
    beta_comm = 0,

    alpha_IS  = qlogis(0.30),
    alpha_pi  = qlogis(0.063),

    sigma_u   = 0.3,
    z_u       = rep(0, stan_data_B$H),

    alpha_phi_I = qlogis(0.85),
    alpha_phi_S = qlogis(0.03),

    mu_raw = c(0, 1),

    sigma_I = 4,
    sigma_S = 3
  )
}

fit_B <- mod_B$sample(
  data = stan_data_B,
  seed = 123,

  chains = 4,
  parallel_chains = 4,     
  threads_per_chain = threads,

  iter_warmup = 2000,
  iter_sampling = 2000,

  init = init_fun_B,
  adapt_delta = 0.97,
  max_treedepth = 12,

  refresh = 50,
  save_warmup = TRUE
)

fit_B



```

## 10. Extract parameters /traceplots and save

```{r}
# ---- Chunk: Extract summaries, LOO/WAIC, and SAVE EVERYTHING (Model B) ----

stopifnot(exists("fit_B"))

# -------------------------------------------------------------------------
# 1) Parameter summaries (key epidemiological parameters)
# -------------------------------------------------------------------------

summary_B <- fit_B$draws(
  variables = c("beta_env","beta_hh","beta_comm","sigma_u",
                "alpha_SI","alpha_IS","alpha_pi")
) %>%
  posterior::as_draws_df() %>%
  select(-.chain, -.draw, -.iteration) %>%
  pivot_longer(everything(), names_to = "parameter", values_to = "value") %>%
  group_by(parameter) %>%
  summarise(
    mean    = mean(value),
    median  = median(value),
    sd      = sd(value),
    q2.5    = quantile(value, 0.025),
    q97.5   = quantile(value, 0.975),
    p_gt_0  = mean(value > 0),
    .groups = "drop"
  ) %>%
  mutate(
    OR_median = if_else(parameter %in% c("beta_env","beta_hh","beta_comm"),
                        exp(median), NA_real_),
    OR_95_lo  = if_else(parameter %in% c("beta_env","beta_hh","beta_comm"),
                        exp(q2.5), NA_real_),
    OR_95_hi  = if_else(parameter %in% c("beta_env","beta_hh","beta_comm"),
                        exp(q97.5), NA_real_)
  ) %>%
  arrange(parameter)

print(summary_B)

# -------------------------------------------------------------------------
# 2) Household-level LOO / WAIC
#    (log_lik_household assumed length H per draw)
# -------------------------------------------------------------------------

loglik_K <- fit_B$draws("log_lik_household") %>%
  posterior::as_draws_matrix()

loo_K  <- loo::loo(loglik_K)
waic_K <- loo::waic(loglik_K)

print(loo_K)
print(waic_K)

# -------------------------------------------------------------------------
# 3) SAVE EVERYTHING I MAY EVER NEED
# -------------------------------------------------------------------------

dir.create(here("outputs", "rds"), recursive = TRUE, showWarnings = FALSE)

# Core reported outputs
saveRDS(summary_B,
        here("outputs", "rds", "summary_B_reducesum_comm.rds"))
saveRDS(loo_K,
        here("outputs", "rds", "loo_K_household.rds"))
saveRDS(waic_K,
        here("outputs", "rds", "waic_K_household.rds"))

# Log-likelihood matrix (for future LOO comparisons)
saveRDS(loglik_K,
        here("outputs", "rds", "loglik_K_household_matrix.rds"))

# Full posterior draws (for plots, new summaries, PPCs)
draws_K <- fit_B$draws() %>% posterior::as_draws_df()
saveRDS(draws_K,
        here("outputs", "rds", "draws_K_df.rds"))

# CmdStan fit object (best insurance against re-running)
fit_B$save_object(
  here("outputs", "rds", "fit_B_cmdstan.rds")
)

# Metadata (sampler info, chains, timing)
saveRDS(fit_B$metadata(),
        here("outputs", "rds", "fit_B_metadata.rds"))

message("Model B fully archived. No re-run required.")


# treaceplot
fit_B$draws(c("beta_hh", "beta_comm", "sigma_u")) |>
  posterior::as_draws_df() |>
  bayesplot::mcmc_trace()

```

\

## 11. Interpretation.

In Model B, we extended a household-level hidden Markov framework to jointly model daily Salmonella acquisition and clearance while explicitly decomposing infection risk into community exposure, household transmission, environmental contamination, and unmeasured household heterogeneity. Incorporating baseline mobility as a proxy for community exposure, we found that the absolute daily probability of acquisition was low at baseline (approximately **2% per day** in the absence of infected household members, environmental contamination, and with average mobility), but increased substantially with social context. Higher community mobility was independently associated with increased risk (odds ratio **1.56**, 95% credible interval **1.03–2.40**), corresponding to an absolute daily risk of approximately **3%** among highly mobile individuals. Within-household transmission exerted the strongest effect, with each additional infected household member more than doubling acquisition risk (odds ratio **2.39**, 95% credible interval **1.72–3.21**), translating to daily risks rising from **\~2%** with no infected household members to **\~5%** with one infected member, **\~11%** with two, and exceeding **20% per day** with three infected members. Baseline household environmental contamination showed weaker but suggestive effects (odds ratio **1.17**, 95% credible interval **0.80–1.70**), while substantial residual between-household variation remained (σ_u ≈ **0.40**). Together, these results indicate that community exposure primarily seeds infection into households, whereas transmission is strongly amplified within households once infection is present, with important implications for interventions—including vaccination—that reduce susceptibility or interrupt short household transmission chains.

## 12. Further save chunk

```{r}
library(digest)
library(here)

# Paths
stan_src  <- here("models", "stan", "model_B_daily_prevstate_reducesum_comm.stan")
out_dir   <- here("outputs", "model_B_archive")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# Copy the exact Stan program used
file.copy(stan_src, file.path(out_dir, basename(stan_src)), overwrite = TRUE)

# Save a fingerprint of the Stan data
setup <- readRDS(here("outputs", "rds", "model_B_setup.rds"))
stan_data_B <- setup$stan_data_B

saveRDS(stan_data_B, file.path(out_dir, "stan_data_B.rds"))
writeLines(digest(stan_data_B, algo = "sha256"), file.path(out_dir, "stan_data_B_sha256.txt"))

# Save session + cmdstan versions
writeLines(capture.output(sessionInfo()), file.path(out_dir, "sessionInfo.txt"))
writeLines(cmdstanr::cmdstan_version(), file.path(out_dir, "cmdstan_version.txt"))

```

```{r}
# ---- Model B: thesis-facing summaries, diagnostics, LOO/WAIC, traceplots ----

library(tidyverse)
library(posterior)
library(ggplot2)
library(here)
library(loo)

# -------------------------------------------------------------------------
# Directories
# -------------------------------------------------------------------------
rds_dir <- here("outputs", "rds")
fig_dir <- here("outputs", "figures")
dir.create(rds_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

# -------------------------------------------------------------------------
# 0) Load saved CmdStanR fit and setup
# -------------------------------------------------------------------------
fit_B <- readRDS(file.path(rds_dir, "fit_B_cmdstan.rds"))
message("Loaded fit_B_cmdstan.rds")

setup_K <- readRDS(file.path(rds_dir, "model_B_setup.rds"))
stan_data_B <- setup_K$stan_data_B
df_long_K   <- setup_K$df_long_K

message("Loaded model_B_setup.rds")

# -------------------------------------------------------------------------
# 1) Try CmdStan diagnostics
# -------------------------------------------------------------------------
csv_files <- tryCatch(
  fit_B$output_files(include_failed = TRUE),
  error = function(e) character()
)

if (length(csv_files) > 0 && all(file.exists(csv_files))) {
  message("CmdStan CSVs found → running cmdstan_diagnose()")
  print(fit_B$cmdstan_diagnose())
} else {
  message(
    "CmdStan CSV files not found or not available.\n",
    "Using draw-based diagnostics instead."
  )
}

# Try sampler diagnostics if available
sampler_diag_K <- tryCatch(
  fit_B$sampler_diagnostics(),
  error = function(e) NULL
)

if (!is.null(sampler_diag_K)) {
  divergences_K <- sum(sampler_diag_K[, , "divergent__"], na.rm = TRUE)
  treedepth_hits_K <- sum(sampler_diag_K[, , "treedepth__"] >= 12, na.rm = TRUE)
} else {
  divergences_K <- NA_integer_
  treedepth_hits_K <- NA_integer_
}

# -------------------------------------------------------------------------
# 2) Extract posterior draws
# -------------------------------------------------------------------------
key_pars <- c(
  "alpha_SI", "alpha_IS", "alpha_pi",
  "beta_env", "beta_hh", "beta_comm", "sigma_u",
  "alpha_phi_I", "alpha_phi_S",
  "mu_I", "mu_S", "sigma_I", "sigma_S"
)

draws_df_K <- tryCatch(
  fit_B$draws(variables = key_pars) |> posterior::as_draws_df(),
  error = function(e) NULL
)

if (is.null(draws_df_K)) {
  message("Falling back to saved draws_K_df.rds")
  draws_df_K <- readRDS(file.path(rds_dir, "draws_K_df.rds")) |>
    posterior::as_draws_df()
}

draws_tbl_K <- draws_df_K |> as_tibble()

missing_key_pars <- setdiff(key_pars, names(draws_tbl_K))
if (length(missing_key_pars) > 0) {
  warning("Missing key parameters from draws: ", paste(missing_key_pars, collapse = ", "))
}

# -------------------------------------------------------------------------
# 3) Draw-based convergence diagnostics
# -------------------------------------------------------------------------
diag_K <- posterior::summarise_draws(
  draws_df_K,
  mean,
  median,
  sd,
  mad,
  q2.5 = ~quantile(.x, probs = 0.025),
  q5   = ~quantile(.x, probs = 0.05),
  q95  = ~quantile(.x, probs = 0.95),
  q97.5 = ~quantile(.x, probs = 0.975),
  rhat,
  ess_bulk,
  ess_tail
)

print(diag_K)

diagnostic_overview_K <- diag_K %>%
  dplyr::filter(variable %in% key_pars) %>%
  dplyr::summarise(
    max_rhat = max(rhat, na.rm = TRUE),
    min_ess_bulk = min(ess_bulk, na.rm = TRUE),
    min_ess_tail = min(ess_tail, na.rm = TRUE),
    divergences = divergences_K,
    treedepth_hits = treedepth_hits_K
  )

print(diagnostic_overview_K)

saveRDS(diag_K, file.path(rds_dir, "diagnostics_K_draw_based.rds"))
saveRDS(diagnostic_overview_K, file.path(rds_dir, "diagnostic_overview_K.rds"))

write_csv(diag_K, file.path(rds_dir, "diagnostics_K_draw_based.csv"))
write_csv(diagnostic_overview_K, file.path(rds_dir, "diagnostic_overview_K.csv"))

# -------------------------------------------------------------------------
# 4) Core posterior parameter summary
# -------------------------------------------------------------------------
parameter_summary_B <- draws_tbl_K %>%
  dplyr::select(any_of(key_pars)) %>%
  tidyr::pivot_longer(
    cols = dplyr::everything(),
    names_to = "parameter",
    values_to = "value"
  ) %>%
  dplyr::group_by(parameter) %>%
  dplyr::summarise(
    mean = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    q025 = quantile(value, 0.025, na.rm = TRUE),
    q05  = quantile(value, 0.05, na.rm = TRUE),
    q95  = quantile(value, 0.95, na.rm = TRUE),
    q975 = quantile(value, 0.975, na.rm = TRUE),
    p_gt_0 = mean(value > 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(parameter)

print(parameter_summary_B)

saveRDS(parameter_summary_B, file.path(rds_dir, "parameter_summary_B_thesis.rds"))
write_csv(parameter_summary_B, file.path(rds_dir, "parameter_summary_B_thesis.csv"))

# -------------------------------------------------------------------------
# 5) Mobility scaling summary
# -------------------------------------------------------------------------
if ("dys_leav_day" %in% names(df_long_K)) {
  mobility_person_K <- df_long_K %>%
    dplyr::distinct(pid, hhid, dys_leav_day, D_star)

  D_bar_K <- mean(mobility_person_K$dys_leav_day, na.rm = TRUE)

  mobility_summary_B <- mobility_person_K %>%
    dplyr::summarise(
      n_people = dplyr::n(),
      mean_days_leaving_house = mean(dys_leav_day, na.rm = TRUE),
      median_days_leaving_house = median(dys_leav_day, na.rm = TRUE),
      min_days_leaving_house = min(dys_leav_day, na.rm = TRUE),
      max_days_leaving_house = max(dys_leav_day, na.rm = TRUE),
      mean_D_star = mean(D_star, na.rm = TRUE),
      median_D_star = median(D_star, na.rm = TRUE),
      min_D_star = min(D_star, na.rm = TRUE),
      max_D_star = max(D_star, na.rm = TRUE)
    )
} else {
  warning("df_long_K does not contain dys_leav_day. Summarising D_star only.")

  mobility_person_K <- purrr::map_dfr(seq_len(stan_data_B$H), function(h) {
    tibble::tibble(
      h = h,
      member_id = seq_len(stan_data_B$M_h[h]),
      D_star = stan_data_B$D_star[h, seq_len(stan_data_B$M_h[h])]
    )
  })

  D_bar_K <- NA_real_

  mobility_summary_B <- mobility_person_K %>%
    dplyr::summarise(
      n_people = dplyr::n(),
      mean_D_star = mean(D_star, na.rm = TRUE),
      median_D_star = median(D_star, na.rm = TRUE),
      min_D_star = min(D_star, na.rm = TRUE),
      max_D_star = max(D_star, na.rm = TRUE)
    )
}

Dstar_never_K <- if (!is.na(D_bar_K)) (0 - D_bar_K) / 7 else NA_real_
Dstar_everyday_K <- if (!is.na(D_bar_K)) (7 - D_bar_K) / 7 else NA_real_

mobility_contrast_K <- tibble::tibble(
  D_bar = D_bar_K,
  D_star_never_leaves_house = Dstar_never_K,
  D_star_leaves_house_every_day = Dstar_everyday_K,
  D_star_difference_every_day_vs_never = Dstar_everyday_K - Dstar_never_K,
  interpretation_beta_comm = "exp(beta_comm) = OR for leaving house every day vs never; exp(beta_comm / 7) = OR per additional day"
)

print(mobility_summary_B)
print(mobility_contrast_K)

saveRDS(mobility_summary_B, file.path(rds_dir, "mobility_summary_B.rds"))
saveRDS(mobility_contrast_K, file.path(rds_dir, "mobility_contrast_K.rds"))

write_csv(mobility_summary_B, file.path(rds_dir, "mobility_summary_B.csv"))
write_csv(mobility_contrast_K, file.path(rds_dir, "mobility_contrast_K.csv"))

# -------------------------------------------------------------------------
# 6) Household data summary
# -------------------------------------------------------------------------
household_tbl_K <- tibble::tibble(
  h = seq_len(stan_data_B$H),
  household_size = stan_data_B$M_h,
  env_positive = stan_data_B$X_h
)

household_model_data_summary_B <- household_tbl_K %>%
  dplyr::summarise(
    n_households = dplyr::n(),
    min_household_size = min(household_size),
    median_household_size = median(household_size),
    max_household_size = max(household_size),
    n_env_positive = sum(env_positive == 1),
    n_env_negative = sum(env_positive == 0),
    prop_env_positive = mean(env_positive == 1)
  )

household_size_table_K <- household_tbl_K %>%
  dplyr::count(household_size, name = "n_households")

print(household_model_data_summary_B)
print(household_size_table_K)

saveRDS(household_model_data_summary_B, file.path(rds_dir, "household_model_data_summary_B.rds"))
saveRDS(household_size_table_K, file.path(rds_dir, "household_size_table_K.rds"))

write_csv(household_model_data_summary_B, file.path(rds_dir, "household_model_data_summary_B.csv"))
write_csv(household_size_table_K, file.path(rds_dir, "household_size_table_K.csv"))

# -------------------------------------------------------------------------
# 7) Derived epidemiological quantities for thesis interpretation
# -------------------------------------------------------------------------
required_for_derived <- c(
  "alpha_SI", "alpha_IS", "alpha_pi",
  "alpha_phi_I", "alpha_phi_S",
  "beta_env", "beta_hh", "beta_comm", "sigma_u",
  "mu_I", "mu_S", "sigma_I", "sigma_S"
)

stopifnot(all(required_for_derived %in% names(draws_tbl_K)))

derived_draws_K <- draws_tbl_K %>%
  dplyr::transmute(
    # Odds ratios
    OR_household_per_infected_member = exp(beta_hh),
    OR_environment_positive_vs_negative = exp(beta_env),
    OR_community_every_day_vs_never = exp(beta_comm),
    OR_community_per_additional_day = exp(beta_comm / 7),
    OR_household_frailty_1SD = exp(sigma_u),

    # Community odds ratios relative to average mobility
    OR_community_every_day_vs_average = exp(beta_comm * Dstar_everyday_K),
    OR_community_never_vs_average = exp(beta_comm * Dstar_never_K),

    # Baseline transition probabilities
    p_acquisition_average_mobility_envneg_no_hh_infected = plogis(alpha_SI),
    p_clearance = plogis(alpha_IS),
    duration_mean_days = 1 / plogis(alpha_IS),
    p_baseline_infected = plogis(alpha_pi),

    # Acquisition probabilities under illustrative scenarios
    p_acquisition_env_positive_avg_mobility_no_hh_infected =
      plogis(alpha_SI + beta_env),

    p_acquisition_never_leaves_envneg_no_hh_infected =
      plogis(alpha_SI + beta_comm * Dstar_never_K),

    p_acquisition_everyday_envneg_no_hh_infected =
      plogis(alpha_SI + beta_comm * Dstar_everyday_K),

    p_acquisition_avg_mobility_one_hh_infected =
      plogis(alpha_SI + beta_hh),

    p_acquisition_avg_mobility_two_hh_infected =
      plogis(alpha_SI + 2 * beta_hh),

    p_acquisition_avg_mobility_three_hh_infected =
      plogis(alpha_SI + 3 * beta_hh),

    # Emission model
    phi_I = plogis(alpha_phi_I),
    phi_S = plogis(alpha_phi_S),
    specificity_like = 1 - plogis(alpha_phi_S),
    mu_I = mu_I,
    mu_S = mu_S,
    sigma_I = sigma_I,
    sigma_S = sigma_S
  )

derived_summary_B <- derived_draws_K %>%
  tidyr::pivot_longer(
    cols = dplyr::everything(),
    names_to = "quantity",
    values_to = "value"
  ) %>%
  dplyr::group_by(quantity) %>%
  dplyr::summarise(
    mean = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    q025 = quantile(value, 0.025, na.rm = TRUE),
    q05  = quantile(value, 0.05, na.rm = TRUE),
    q95  = quantile(value, 0.95, na.rm = TRUE),
    q975 = quantile(value, 0.975, na.rm = TRUE),
    p_gt_1 = mean(value > 1, na.rm = TRUE),
    p_gt_0 = mean(value > 0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(quantity)

print(derived_summary_B)

saveRDS(derived_summary_B, file.path(rds_dir, "derived_summary_B_thesis.rds"))
write_csv(derived_summary_B, file.path(rds_dir, "derived_summary_B_thesis.csv"))

# A compact table of the main odds ratios for the chapter
main_OR_table_K <- derived_summary_B %>%
  dplyr::filter(
    quantity %in% c(
      "OR_household_per_infected_member",
      "OR_community_every_day_vs_never",
      "OR_community_per_additional_day",
      "OR_environment_positive_vs_negative",
      "OR_household_frailty_1SD"
    )
  ) %>%
  dplyr::select(quantity, median, q025, q975)

print(main_OR_table_K)

saveRDS(main_OR_table_K, file.path(rds_dir, "main_OR_table_K_thesis.rds"))
write_csv(main_OR_table_K, file.path(rds_dir, "main_OR_table_K_thesis.csv"))

# Compact absolute-risk scenario table
absolute_risk_table_K <- derived_summary_B %>%
  dplyr::filter(
    quantity %in% c(
      "p_acquisition_average_mobility_envneg_no_hh_infected",
      "p_acquisition_env_positive_avg_mobility_no_hh_infected",
      "p_acquisition_never_leaves_envneg_no_hh_infected",
      "p_acquisition_everyday_envneg_no_hh_infected",
      "p_acquisition_avg_mobility_one_hh_infected",
      "p_acquisition_avg_mobility_two_hh_infected",
      "p_acquisition_avg_mobility_three_hh_infected",
      "p_clearance",
      "duration_mean_days",
      "p_baseline_infected"
    )
  ) %>%
  dplyr::select(quantity, median, q025, q975)

print(absolute_risk_table_K)

saveRDS(absolute_risk_table_K, file.path(rds_dir, "absolute_risk_table_K_thesis.rds"))
write_csv(absolute_risk_table_K, file.path(rds_dir, "absolute_risk_table_K_thesis.csv"))

# -------------------------------------------------------------------------
# 8) LOO / WAIC and Pareto-k diagnostics
# -------------------------------------------------------------------------
loglik_K <- tryCatch(
  fit_B$draws("log_lik_household") |> posterior::as_draws_matrix(),
  error = function(e) NULL
)

if (is.null(loglik_K) && file.exists(file.path(rds_dir, "loglik_K_household_matrix.rds"))) {
  message("Falling back to saved loglik_K_household_matrix.rds")
  loglik_K <- readRDS(file.path(rds_dir, "loglik_K_household_matrix.rds"))
}

if (!is.null(loglik_K)) {
  loo_K  <- loo::loo(loglik_K)
  waic_K <- loo::waic(loglik_K)

  print(loo_K)
  print(waic_K)

  pareto_K <- tibble::tibble(
    household_index = seq_along(loo::pareto_k_values(loo_K)),
    pareto_k = loo::pareto_k_values(loo_K)
  ) %>%
    dplyr::mutate(
      k_group = dplyr::case_when(
        pareto_k <= 0.5 ~ "<=0.5",
        pareto_k <= 0.7 ~ "0.5-0.7",
        pareto_k <= 1.0 ~ "0.7-1.0",
        TRUE ~ ">1.0"
      )
    )

  pareto_summary_B <- pareto_K %>%
    dplyr::count(k_group, name = "n_households") %>%
    dplyr::arrange(k_group)

  pareto_top_K <- pareto_K %>%
    dplyr::arrange(desc(pareto_k)) %>%
    dplyr::slice_head(n = 10)

  print(pareto_summary_B)
  print(pareto_top_K)

  saveRDS(loo_K, file.path(rds_dir, "loo_K_household.rds"))
  saveRDS(waic_K, file.path(rds_dir, "waic_K_household.rds"))
  saveRDS(loglik_K, file.path(rds_dir, "loglik_K_household_matrix.rds"))
  saveRDS(pareto_K, file.path(rds_dir, "pareto_K_household.rds"))
  saveRDS(pareto_summary_B, file.path(rds_dir, "pareto_summary_B_household.rds"))
  saveRDS(pareto_top_K, file.path(rds_dir, "pareto_top_K_household.rds"))

  write_csv(pareto_K, file.path(rds_dir, "pareto_K_household.csv"))
  write_csv(pareto_summary_B, file.path(rds_dir, "pareto_summary_B_household.csv"))
  write_csv(pareto_top_K, file.path(rds_dir, "pareto_top_K_household.csv"))
} else {
  message("log_lik_household not available → skipping LOO/WAIC")
}

# Optional: compare Model J vs Model B if Model J household log-likelihood exists
if (file.exists(file.path(rds_dir, "loglik_J_household_matrix.rds")) && !is.null(loglik_K)) {
  loglik_J <- readRDS(file.path(rds_dir, "loglik_J_household_matrix.rds"))

  if (ncol(loglik_J) == ncol(loglik_K)) {
    loo_J <- loo::loo(loglik_J)
    loo_compare_JK <- loo::loo_compare(loo_J, loo_K)

    print(loo_compare_JK)

    saveRDS(loo_compare_JK, file.path(rds_dir, "loo_compare_J_vs_K.rds"))
    write_csv(
      as.data.frame(loo_compare_JK) %>% tibble::rownames_to_column("model"),
      file.path(rds_dir, "loo_compare_J_vs_K.csv")
    )
  } else {
    message("Model J and Model B log-likelihood matrices have different numbers of columns; skipping comparison.")
  }
}

# -------------------------------------------------------------------------
# 9) Traceplots
# -------------------------------------------------------------------------
draws_long <- draws_tbl_K %>%
  dplyr::select(.chain, .iteration, .draw, any_of(key_pars)) %>%
  tidyr::pivot_longer(
    cols = any_of(key_pars),
    names_to = "parameter",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    parameter = factor(parameter, levels = key_pars),
    chain = factor(.chain)
  )

p_trace <- ggplot(draws_long, aes(x = .iteration, y = value, colour = chain)) +
  geom_line(alpha = 0.6, linewidth = 0.3) +
  facet_wrap(~parameter, scales = "free_y") +
  labs(
    x = "Iteration",
    y = "Value",
    colour = "Chain",
    title = "Traceplots for key Model B parameters"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")

p_trace

ggsave(
  filename = file.path(fig_dir, "trace_ModelK_key_parameters.png"),
  plot = p_trace,
  dpi = 300,
  width = 12,
  height = 7,
  bg = "white"
)

# -------------------------------------------------------------------------
# 10) Thesis-facing forest plot
# -------------------------------------------------------------------------
effects_sum_K <- derived_draws_K %>%
  dplyr::transmute(
    `Household amplification\n(per additional infected household member yesterday)` =
      OR_household_per_infected_member,
    `Community exposure\n(leaving house every day vs never)` =
      OR_community_every_day_vs_never,
    `Community exposure\n(per additional day leaving the house)` =
      OR_community_per_additional_day,
    `Household environment\n(environmental contamination positive vs negative)` =
      OR_environment_positive_vs_negative
  ) %>%
  tidyr::pivot_longer(
    cols = dplyr::everything(),
    names_to = "Effect",
    values_to = "OR"
  ) %>%
  dplyr::group_by(Effect) %>%
  dplyr::summarise(
    OR_med = median(OR, na.rm = TRUE),
    lo = quantile(OR, 0.025, na.rm = TRUE),
    hi = quantile(OR, 0.975, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    label = sprintf("%.2f [%.2f to %.2f]", OR_med, lo, hi),
    Effect = factor(
      Effect,
      levels = c(
        "Household amplification\n(per additional infected household member yesterday)",
        "Community exposure\n(leaving house every day vs never)",
        "Community exposure\n(per additional day leaving the house)",
        "Household environment\n(environmental contamination positive vs negative)"
      )
    )
  )

x_min <- min(effects_sum_K$lo, na.rm = TRUE) * 0.85
x_max <- max(effects_sum_K$hi, na.rm = TRUE) * 1.20

p_forest_K <- ggplot(effects_sum_K, aes(y = Effect, x = OR_med)) +
  geom_vline(xintercept = 1, linetype = "22", linewidth = 0.6, colour = "grey40") +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.22, linewidth = 1.0) +
  geom_point(size = 3.3) +
  geom_text(
    aes(x = pmin(hi * 1.08, x_max), label = label),
    hjust = 0,
    size = 3.5
  ) +
  scale_x_log10(
    breaks = c(0.5, 0.75, 1, 1.25, 1.5, 2, 3, 4),
    labels = c("0.5", "0.75", "1", "1.25", "1.5", "2", "3", "4"),
    limits = c(x_min, x_max)
  ) +
  labs(
    x = "Odds ratio (log scale)",
    y = NULL,
    title = "Drivers of next-day acquisition in the coupled household model",
    subtitle = "Posterior median odds ratios with 95% credible intervals"
  ) +
  theme_classic(base_size = 13)

p_forest_K

ggsave(
  filename = file.path(fig_dir, "forest_ModelK_effects_thesis.png"),
  plot = p_forest_K,
  dpi = 400,
  width = 12,
  height = 5.5,
  bg = "white"
)

# -------------------------------------------------------------------------
# 11) Final message
# -------------------------------------------------------------------------
message("Model B thesis summaries complete.")
message("Key files saved in: ", rds_dir)
message("Key figures saved in: ", fig_dir)

### now ouput all summaries:
purrr::iwalk(setNames(purrr::map(c("parameter_summary_B_thesis.csv","derived_summary_B_thesis.csv","main_OR_table_K_thesis.csv","absolute_risk_table_K_thesis.csv","mobility_summary_B.csv","mobility_contrast_K.csv","household_model_data_summary_B.csv","household_size_table_K.csv","pareto_summary_B_household.csv","diagnostic_overview_K.csv"), ~ readr::read_csv(here::here("outputs","rds", .x), show_col_types = FALSE)), c("parameter_summary_B","derived_summary_B","main_OR_table_K","absolute_risk_table_K","mobility_summary_B","mobility_contrast_K","household_model_data_summary_B","household_size_table_K","pareto_summary_B","diagnostic_overview_K")), ~ {cat("\n\n==================== ", .y, " ====================\n", sep = ""); print(.x, n = Inf, width = Inf)}) 
```

```{r}
# ---- Publication-ready forest plot: intuitive labels + clean styling (no refit) ----
# Outputs: outputs/figures/forest_modelK_effects_pubready.png

library(tidyverse)
library(posterior)
library(ggplot2)
library(scales)
library(here)

# Paths
rds_path <- here("outputs", "rds", "draws_K_df.rds")
fig_dir  <- here("outputs", "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(file.exists(rds_path))

# Load draws (NO rerun)
draws <- readRDS(rds_path) |>
  posterior::as_draws_df() |>
  as_tibble()

# Intuitive effect scales:
# - beta_hh: OR per additional infected household member yesterday
# - beta_env: OR for environmentally positive vs negative household
# - beta_comm: D_star = (days_out - mean)/7, so:
#     OR per +1 day out/week = exp(beta_comm/7)
#     OR for "7 days out vs 0 days out" = exp(beta_comm)
effects_sum <- draws |>
  transmute(
    `Household amplification\n(per additional infected household member yesterday)` = exp(beta_hh),
    `Community exposure\n(per additional day leaving the house in last week)`      = exp(beta_comm / 7),
    `Community exposure\n(leaving house every day vs never)`                       = exp(beta_comm),
    `Household environment\n(environmental contamination positive vs negative)`    = exp(beta_env)
  ) |>
  pivot_longer(everything(), names_to = "Effect", values_to = "OR") |>
  group_by(Effect) |>
  summarise(
    OR_med = median(OR),
    lo     = quantile(OR, 0.025),
    hi     = quantile(OR, 0.975),
    .groups = "drop"
  ) |>
  mutate(
    # Human-readable inline label for the plot
    label = sprintf("%.2f  [%.2f to %.2f]", OR_med, lo, hi),
    # Order effects so the eye lands on the story: household >> community >> environment
    Effect = factor(
      Effect,
      levels = c(
        "Household amplification\n(per additional infected household member yesterday)",
        "Community exposure\n(leaving house every day vs never)",
        "Community exposure\n(per additional day leaving the house in last week)",
        "Household environment\n(environmental contamination positive vs negative)"
      )
    )
  ) |>
  arrange(Effect)

print(effects_sum)

# Choose tidy log-scale breaks that read well on slides/paper
x_breaks <- c(0.5, 0.75, 1, 1.25, 1.5, 2, 3, 4)

# Add a little padding so CI bars do not hit the edges
x_min <- min(effects_sum$lo, na.rm = TRUE) * 0.85
x_max <- max(effects_sum$hi, na.rm = TRUE) * 1.15

p_forest <- ggplot(effects_sum, aes(y = Effect, x = OR_med)) +
  # Reference line at no effect
  geom_vline(xintercept = 1, linetype = "22", linewidth = 0.6, colour = "grey35") +
  # Credible intervals
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.20, linewidth = 0.9, colour = "grey15") +
  # Point estimate
  geom_point(size = 3.2, shape = 21, stroke = 0.9, fill = "white", colour = "grey10") +
  # Text label with OR and interval (placed to the right of the CI)
  geom_text(
    aes(x = pmin(hi * 1.08, x_max), label = label),
    hjust = 0, size = 3.6, colour = "grey10"
  ) +
  scale_x_log10(
    breaks = x_breaks,
    labels = x_breaks,
    limits = c(x_min, x_max)
  ) +
  labs(
    title = "What drives next-day acquisition in the coupled household model?",
    subtitle = "Posterior odds ratios (median and 95% credible interval). Household amplification dominates.",
    x = "Odds ratio (log scale)",
    y = NULL,
    caption = "Community exposure is measured once at baseline (days leaving the house in the previous week)."
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title.position = "plot",
    axis.text.y = element_text(size = 12.5),
    axis.text.x = element_text(size = 11.5),
    axis.title.x = element_text(size = 12.5, margin = margin(t = 8)),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 10.5, colour = "grey25", margin = margin(t = 10)),
    axis.ticks.y = element_blank()
  )

p_forest

# Save high-res for slides/paper
out_png <- file.path(fig_dir, "forest_modelK_effects_pubready.png")
ggsave(out_png, plot = p_forest, width = 11.5, height = 5.2, dpi = 400, bg = "white")
message("Saved: ", out_png)


```

```{r}
# ---- Publication-ready forest plot with semantic colour (no refit) ----

library(tidyverse)
library(posterior)
library(ggplot2)
library(here)

# Paths
rds_path <- here("outputs", "rds", "draws_K_df.rds")
fig_dir  <- here("outputs", "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(file.exists(rds_path))

# Load posterior draws
draws <- readRDS(rds_path) |>
  posterior::as_draws_df() |>
  as_tibble()

# Summarise effects on intuitive scales
effects_sum <- draws |>
  transmute(
    Effect = c(
      "Household amplification\n(per additional infected household member yesterday)",
      "Community exposure\n(leaving house every day vs never)",
      "Community exposure\n(per additional day leaving the house)",
      "Household environment\n(environmental contamination positive vs negative)"
    ),
    OR = list(
      exp(beta_hh),
      exp(beta_comm),
      exp(beta_comm / 7),
      exp(beta_env)
    )
  ) |>
  unnest(OR) |>
  group_by(Effect) |>
  summarise(
    OR_med = median(OR),
    lo     = quantile(OR, 0.025),
    hi     = quantile(OR, 0.975),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("%.2f  [%.2f to %.2f]", OR_med, lo, hi),
    mechanism = case_when(
      str_detect(Effect, "Household amplification") ~ "Household transmission",
      str_detect(Effect, "Community exposure")      ~ "Community exposure",
      TRUE                                          ~ "Environment"
    ),
    Effect = factor(
      Effect,
      levels = c(
        "Household amplification\n(per additional infected household member yesterday)",
        "Community exposure\n(leaving house every day vs never)",
        "Community exposure\n(per additional day leaving the house)",
        "Household environment\n(environmental contamination positive vs negative)"
      )
    )
  )

# Colour palette (Okabe–Ito inspired; colour-blind safe)
cols <- c(
  "Household transmission" = "#0072B2",  # blue
  "Community exposure"     = "#D55E00",  # orange
  "Environment"            = "#666666"   # grey
)

# Axis limits with padding
x_min <- min(effects_sum$lo) * 0.85
x_max <- max(effects_sum$hi) * 1.15

p_forest <- ggplot(effects_sum, aes(y = Effect, x = OR_med, colour = mechanism)) +
  # No-effect line
  geom_vline(xintercept = 1, linetype = "22", linewidth = 0.6, colour = "grey40") +
  # Credible intervals
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.22, linewidth = 1.0) +
  # Point estimates
  geom_point(size = 3.6, shape = 16) +
  # Numeric labels
  geom_text(
    aes(x = pmin(hi * 1.08, x_max), label = label),
    hjust = 0, size = 3.7, colour = "black"
  ) +
  scale_colour_manual(values = cols) +
  scale_x_log10(
    breaks = c(0.5, 0.75, 1, 1.25, 1.5, 2, 3, 4),
    labels = c("0.5","0.75","1","1.25","1.5","2","3","4"),
    limits = c(x_min, x_max)
  ) +
  labs(
    title = "What drives next-day Salmonella acquisition within households?",
    subtitle = "Posterior odds ratios from the coupled household model (median and 95% credible intervals)",
    x = "Odds ratio (log scale)",
    y = NULL,
    colour = NULL,
    caption = "Community exposure is measured at baseline (days leaving the house in the prior week)."
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12.5, margin = margin(b = 10)),
    axis.text.y = element_text(size = 12.5),
    axis.text.x = element_text(size = 11.5),
    axis.title.x = element_text(size = 12.5, margin = margin(t = 8)),
    legend.position = "top",
    plot.caption = element_text(size = 10.5, colour = "grey30", margin = margin(t = 10))
  )

p_forest

# Save high-resolution output
out_png <- file.path(fig_dir, "forest_modelK_effects_colour_pubready.png")
ggsave(out_png, plot = p_forest, width = 11.5, height = 5.3, dpi = 400, bg = "white")
message("Saved: ", out_png)

```

```{r}

# ---- Publication-ready forest plot: coupled household model covariate effects ----

library(tidyverse)
library(posterior)
library(ggplot2)
library(here)

# Paths
rds_path <- here("outputs", "rds", "draws_K_df.rds")
fig_dir  <- here("outputs", "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(file.exists(rds_path))

# Load posterior draws
draws <- readRDS(rds_path) |>
  posterior::as_draws_df() |>
  as_tibble()

# Check required posterior parameters are present
required_params <- c("beta_hh", "beta_comm", "beta_env")
missing_params <- setdiff(required_params, names(draws))

if (length(missing_params) > 0) {
  stop(
    "Missing required posterior parameter(s): ",
    paste(missing_params, collapse = ", ")
  )
}

# -------------------------------------------------------------------------
# Assumption:
# beta_comm represents the model coefficient for leaving the house every day
# versus never leaving the house across the 7-day period.
#
# If beta_comm was instead fitted to the raw number of days out, then the
# 7 days vs 0 days contrast should be exp(beta_comm * 7), not exp(beta_comm).
# -------------------------------------------------------------------------

effects_long <- draws |>
  transmute(
    `Household infection burden\nper additional infected member` =
      exp(beta_hh),

    `Mobility outside household\n7 days vs 0 days` =
      exp(beta_comm),

    `Baseline household contamination\npositive vs negative` =
      exp(beta_env)
  ) |>
  pivot_longer(
    cols = everything(),
    names_to = "Effect",
    values_to = "OR"
  )

# Desired visual order: top to bottom
plot_order <- c(
  "Household infection burden\nper additional infected member",
  "Mobility outside household\n7 days vs 0 days",
  "Baseline household contamination\npositive vs negative"
)

# Summarise posterior
effects_sum <- effects_long |>
  group_by(Effect) |>
  summarise(
    OR_med = median(OR),
    lo     = quantile(OR, 0.025),
    hi     = quantile(OR, 0.975),
    .groups = "drop"
  ) |>
  mutate(
    estimate_label = sprintf("%.2f [%.2f to %.2f]", OR_med, lo, hi),

    # ggplot places the first y-level at the bottom,
    # so reverse the levels to get the desired top-to-bottom order.
    Effect = factor(Effect, levels = rev(plot_order))
  )

# Muted Lancet-style colours
# These are taken from the Lancet-style palette family but avoid a primary-colour look.
lancet_cols <- c(
  "Household infection burden\nper additional infected member" = "#00468B",
  "Mobility outside household\n7 days vs 0 days"               = "#0099B4",
  "Baseline household contamination\npositive vs negative"     = "#925E9F"
)

# Log-scale axis limits with room for right-hand estimate labels
x_values <- c(effects_sum$lo, effects_sum$hi, 1)

log_range <- range(log10(x_values), na.rm = TRUE)
pad <- diff(log_range) * 0.15

if (!is.finite(pad) || pad == 0) {
  pad <- 0.15
}

x_left  <- 10^(log_range[1] - pad)

# Position of numerical labels
x_text  <- 10^(log_range[2] + pad * 1.15)

# Keep the plotting area elegant, but allow space for labels using clip = "off"
x_limit <- 10^(log_range[2] + pad * 2.70)

x_breaks <- c(1, 2, 4)
x_breaks <- x_breaks[x_breaks >= x_left & x_breaks <= x_limit]

# Plot
p_forest <- ggplot(
  effects_sum,
  aes(y = Effect, x = OR_med)
) +
  geom_vline(
    xintercept = 1,
    linetype = "22",
    linewidth = 0.65,
    colour = "grey45"
  ) +

  # Credible interval bars
  geom_segment(
    aes(
      x = lo,
      xend = hi,
      y = Effect,
      yend = Effect,
      colour = Effect
    ),
    linewidth = 1.85,
    lineend = "round"
  ) +

  # Posterior median points
  geom_point(
    aes(fill = Effect),
    shape = 21,
    size = 5.7,
    colour = "white",
    stroke = 0.95
  ) +

  # Right-hand posterior summaries
  geom_text(
    data = effects_sum,
    aes(
      y = Effect,
      x = x_text,
      label = estimate_label
    ),
    inherit.aes = FALSE,
    hjust = 0,
    size = 4.45,
    colour = "black"
  ) +

  scale_colour_manual(values = lancet_cols) +
  scale_fill_manual(values = lancet_cols) +

  scale_x_log10(
    breaks = x_breaks,
    limits = c(x_left, x_limit),
    labels = function(x) format(x, trim = TRUE, scientific = FALSE)
  ) +

  coord_cartesian(clip = "off") +

  labs(
    x = "Posterior odds ratio for acquisition",
    y = NULL
  ) +

  theme_classic(base_size = 15.5) +
  theme(
    legend.position = "none",

    axis.line.y  = element_blank(),
    axis.ticks.y = element_blank(),

    axis.line.x = element_line(
      linewidth = 0.65,
      colour = "black"
    ),

    axis.ticks.x = element_line(
      linewidth = 0.65,
      colour = "black"
    ),

    axis.text.y = element_text(
      size = 16.5,
      lineheight = 1.10,
      colour = "black"
    ),

    axis.text.x = element_text(
      size = 14.5,
      colour = "black"
    ),

    axis.title.x = element_text(
      size = 16.5,
      margin = margin(t = 11)
    ),

    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),

    plot.margin = margin(10, 215, 10, 10)
  )

p_forest

# Save source data
write_csv(
  effects_sum |>
    mutate(
      Effect = as.character(Effect),
      statistic = "posterior median and 95% credible interval"
    ),
  file.path(fig_dir, "forest_modelK_effects_source_data.csv")
)

# Save figure
out_png <- file.path(fig_dir, "forest_modelK_effects_lancet_colour.png")
out_pdf <- file.path(fig_dir, "forest_modelK_effects_lancet_colour.pdf")

ggsave(
  out_png,
  plot = p_forest,
  width = 14.7,
  height = 5.25,
  dpi = 500,
  bg = "white"
)

pdf_device <- if (capabilities("cairo")) cairo_pdf else "pdf"

ggsave(
  out_pdf,
  plot = p_forest,
  device = pdf_device,
  width = 14.7,
  height = 5.25,
  bg = "white"
)

message("Saved: ", out_png)
message("Saved: ", out_pdf)




```

## 13. Plotting chunk of within-house acquisition pressure

```{r}
# ---- Figure 1: acquisition probability by household infection pressure ----
# No annotation; clearer percentage labels
# Output: outputs/figures/Figure1_acquisition_probability_household_pressure_clean.png

library(tidyverse)
library(posterior)
library(ggplot2)
library(here)
library(scales)

fig_dir <- here("outputs", "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

# Load posterior draws
draws_K <- readRDS(here("outputs", "rds", "draws_K_df.rds")) |>
  posterior::as_draws_df() |>
  as_tibble()

stopifnot(all(c("alpha_SI", "beta_hh") %in% names(draws_K)))

# Posterior-predicted acquisition probabilities
# Scenario: environmentally negative household, reference mobility, household frailty = 0
acq_pressure_draws <- draws_K |>
  dplyr::select(alpha_SI, beta_hh) |>
  tidyr::crossing(n_infected_previous = 0:3) |>
  dplyr::mutate(
    p_acquisition = plogis(alpha_SI + beta_hh * n_infected_previous)
  )

acq_pressure_summary <- acq_pressure_draws |>
  group_by(n_infected_previous) |>
  summarise(
    median = median(p_acquisition, na.rm = TRUE),
    lo     = quantile(p_acquisition, 0.025, na.rm = TRUE),
    hi     = quantile(p_acquisition, 0.975, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    median_pct = 100 * median,
    lo_pct     = 100 * lo,
    hi_pct     = 100 * hi,
    label      = sprintf("%.1f%%", median_pct),
    label_y    = case_when(
      n_infected_previous == 0 ~ median_pct + 2.1,
      n_infected_previous == 1 ~ median_pct + 2.1,
      n_infected_previous == 2 ~ median_pct + 2.6,
      n_infected_previous == 3 ~ median_pct + 3.6
    )
  )

print(acq_pressure_summary)

write_csv(
  acq_pressure_summary,
  here("outputs", "rds", "Figure1_acquisition_probability_household_pressure_clean.csv")
)

# Interpolated curve for a smooth visual connection
curve_df <- tibble(
  n_infected_previous = seq(0, 3, length.out = 200)
) |>
  mutate(
    median_pct = approx(
      x = acq_pressure_summary$n_infected_previous,
      y = acq_pressure_summary$median_pct,
      xout = n_infected_previous
    )$y,
    lo_pct = approx(
      x = acq_pressure_summary$n_infected_previous,
      y = acq_pressure_summary$lo_pct,
      xout = n_infected_previous
    )$y,
    hi_pct = approx(
      x = acq_pressure_summary$n_infected_previous,
      y = acq_pressure_summary$hi_pct,
      xout = n_infected_previous
    )$y
  )

blue_main <- "#005EA8"
blue_fill <- "#A6CEE3"
grey_text <- "#202020"

p_fig1_clean <- ggplot() +
  geom_ribbon(
    data = curve_df,
    aes(
      x = n_infected_previous,
      ymin = lo_pct,
      ymax = hi_pct
    ),
    fill = blue_fill,
    alpha = 0.35
  ) +
  geom_line(
    data = curve_df,
    aes(x = n_infected_previous, y = median_pct),
    colour = blue_main,
    linewidth = 1.2,
    lineend = "round"
  ) +
  geom_errorbar(
    data = acq_pressure_summary,
    aes(
      x = n_infected_previous,
      ymin = lo_pct,
      ymax = hi_pct
    ),
    width = 0.07,
    linewidth = 0.85,
    colour = blue_main
  ) +
  geom_point(
    data = acq_pressure_summary,
    aes(x = n_infected_previous, y = median_pct),
    size = 4.0,
    colour = blue_main,
    fill = "white",
    shape = 21,
    stroke = 1.25
  ) +
  geom_label(
    data = acq_pressure_summary,
    aes(
      x = n_infected_previous,
      y = label_y,
      label = label
    ),
    size = 4.4,
    fontface = "bold",
    colour = grey_text,
    fill = "white",
    label.size = 0,
    label.padding = unit(0.13, "lines")
  ) +
  scale_x_continuous(
    breaks = 0:3,
    limits = c(-0.25, 3.25),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_continuous(
    breaks = seq(0, 45, by = 5),
    limits = c(0, 45),
    labels = label_number(suffix = "%"),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title = "Household infection pressure strongly increases acquisition",
    subtitle = "Posterior-predicted daily acquisition probability in an environmentally negative household at reference mobility",
    x = "Latently infected household members at previous time step",
    y = "Daily probability of acquisition",
    caption = "Points show posterior medians; vertical bars and shaded band show 95% credible intervals."
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title.position = "plot",
    plot.caption.position = "plot",

    plot.title = element_text(
      face = "bold",
      size = 17,
      colour = "black",
      margin = margin(b = 6)
    ),
    plot.subtitle = element_text(
      size = 11.5,
      colour = "grey25",
      margin = margin(b = 16)
    ),
    plot.caption = element_text(
      size = 9.5,
      colour = "grey35",
      margin = margin(t = 12)
    ),

    axis.title.x = element_text(
      size = 12,
      colour = "black",
      margin = margin(t = 10)
    ),
    axis.title.y = element_text(
      size = 12,
      colour = "black",
      margin = margin(r = 10)
    ),
    axis.text = element_text(
      size = 11,
      colour = "black"
    ),

    axis.line = element_line(linewidth = 0.55, colour = "black"),
    axis.ticks = element_line(linewidth = 0.45, colour = "black"),
    axis.ticks.length = unit(0.18, "cm"),

    panel.grid.major.y = element_line(
      colour = "grey88",
      linewidth = 0.35
    ),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),

    plot.margin = margin(t = 14, r = 24, b = 12, l = 14)
  )

p_fig1_clean

out_png <- here(
  "outputs",
  "figures",
  "Figure1_acquisition_probability_household_pressure_clean.png"
)

ggsave(
  filename = out_png,
  plot = p_fig1_clean,
  width = 8.8,
  height = 5.8,
  dpi = 300,
  bg = "white"
)

message("Saved: ", out_png)
```
